---
title       : Working across the data
description : Cross-tabulations let you explore dependencies hidden deep within discrete variables.

--- type:NormalExercise lang:r xp:100 skills:1 key:e1b927030f

## The sum of the variable

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:24b2372fff

## Tables

table() & prop.table()

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:6fad9de433

## Metadata of students2014

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:a7f8986e45

## Cross tabulations

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:242f06b715

## Let's plot that table

mosaicplot()/plot()
barplot(..., beside=T/F)

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

