---
title       : Describing and exploring variability
description : Variability is at the heart of statistics. In fact, without it the discipline would seize to exist. R offers great tools for exploring the variability in your data.

--- type:NormalExercise lang:r xp:100 skills:1, 3
## summary() statistics

Some functions in R are very generic. They can take different objects as their first argument and they seem to magically do exactly what you'd hope they do in all situations.

One such example is the `summary()` function.

*** =instructions
- Execute the code that creates a summary of the `variable` in `df`
- Then create a summary of `df`
- `summary()` behaves differently because `variable` and `df` are different kinds of objects.

*** =hint
- Get the summary of an object by typing summary(object)

*** =pre_exercise_code
```{r}
# pec
```


*** =sample_code
```{r}

# Create a summary of var
summary(df$var)

# Create a summary of df



```

*** =solution
```{r}

# Create a summary of var in df
summary(df$var)

# Create a summary of df
summary(df)


```

*** =sct
```{r}
# tests. 
test_output_contains("summary(df)", incorrect_msg="Please type in the code to create a summary of `df`")

test_error()
success_msg("Great job!")
```

--- type:NormalExercise lang:r xp:100 skills:1

## Help!

How to use help() and ? (and str() )

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1

## Bar plots

Bar plots are useful for visualizing qualitative variables, which in R are usually stored as character-, factor- or boolean data types.

In base R graphics, the function `barplot()` can be used to draw a bar plot. It's first argument is `height` which is the only mandatory argument. The information for `height` can be computed easily by creating a `summary()` or a `table()` of a factor.

In this exercise you will draw a barplot of `variable` in the data frame `df`. You can use `str()` to see the structure of `df`.

*** =instructions
- Remind yourself on how to access the column of a data frame and how to create objects.
- Use `table()` on `variable` in the data frame `df`. What does it show you?
- Create an object `counts` by assigning the output of `table()` or `summary()` to it.
- Use the function `barplot()` to draw a bar plot. Use the object you created as the first argument.

*** =hint
- You can access the columns of a data.frame (df) with the `$` operator: `df$column`.
- Remember how to create objects using the assign operator `<-`: `object_name <- function()`


*** =pre_exercise_code
```{r}
# pec
```


*** =sample_code
```{r}
# Look at the counts of different levels.
table(df)


# Create object counts that stores the counts
counts <- NULL


# Call barplot() to draw a graph. Use the object you created as an argument.

  
```

*** =solution
```{r}
# Look at the counts of different levels.
table()


# Create object counts that stores the counts
counts <- table()


# Call barplot() to draw a graph. Use the object you created as an argument.
barplot(counts)



```

*** =sct
```{r}

test_function("barplot", args="height")

# test error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Excellent work!")

```

--- type:NormalExercise lang:r xp:100 skills:1

## Histograms

Histograms are an excellent way of visualizing the variability of a continuous quantitative variable. Basic histograms can be drawn with the `hist()` function. The first argument of `hist()`, `x` should be a numeric vector of data values.

As with all the plotting functions in R, the `hist()` function can take on many additional arguments which allow for more control over the plot. One simple argument is `col`, which defines the color of the bars.

`col` can be a character string giving the name of the color ("blue", "green", etc). Some possible color names can be found <a target="_blank" href="http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf">here</a> The `breaks` argument can be used to control the approximate number of bins in the histogram.

*** =instructions
- Create a histogram of 
- Adjust your code and draw the histogram again using `breaks = 12`.
- Adjust your code and give your histogram a nice color using the `col` argument. Use quotations marks to define the color name.


*** =hint
- You can access the column of a data frame (df) with the `$` symbol: `df$column_name`.
- Remember to seperate the arguments of a function with a comma.


*** =pre_exercise_code
```{r}
# pec
```


*** =sample_code
```{r}


# Draw a histogram of var. Use the additional arguments breaks and col.

histogram(var, col = NULL)




```

*** =solution
```{r}
# Draw a histogram of Income in US. Use the additional arguments breaks and col.
hist(var, breaks = 12, col ="blue")


```

*** =sct
```{r}

test_function("hist", args=c("x","breaks"))

# test error
test_error()


# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:150 skills:1

## Box plots

One very good way to visualize the most important summary statistics of a variable is by drawing box plots. A box plot visualizes the 25th, 50th and 75th percentiles (the box), the typical range (the whiskers) and the outliers (the dots) of a variable 

The whiskers extending from the box can be computed by several techniques. The default in R is to extend them to reach to a data point that is no more than 1.5*IQR away from the box, where IQR is the inter quartile range defined as  

`IQR = 75th percentile - 25th percentile`  

Values outside the whiskers can be considered as outliers, unusually distant observations. For more information on IQR, see <a target="_blank" href ="https://en.wikipedia.org/wiki/Interquartile_range"> wikipedia</a>

*** =instructions
- Draw a `boxplot()` of `var`.
- Use help() or `?` to find out which argument controls the label of the y axis and the main title.
- Add a suitable label to the y axis and a main title of your choice.

*** =hint
- hint1

*** =pre_exercise_code
```{r}
# pec
```


*** =sample_code
```{r}
# data frame is available

# See the help page of boxplot()
?boxplot

# Draw a boxplot of var. Add a y axis label and a main title.



```

*** =solution
```{r}
# Draw a boxplot of var. Add a y axis label and a main title.
boxplot(var, ylab = "", main = "")


```

*** =sct
```{r}


test_function("boxplot", args="x")

# test error
test_error()


# Final message the student will see upon completing the exercise
success_msg("Awsome work!")

```

