---
title       : Confidence
description : UNDER CONSTRUCTION


--- type:NormalExercise lang:r xp:100 skills:1

## The confidence interval

Intriduction to the subject of the sampling distribution and confidence.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# exampe code

```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:23abc9f934

## Indeces and brackets

Vectors store multiple values of the same data type. The values in vectors have indices: the first value in a vector has a index `1`, the second value has `2` and so on. 

You can set or get a single value from a vector by using indices and bracets `[ ]`. Using an index number inside the bracets gives access to a single value from the vector. Using a vector of indices gives access to a vector of values.

It is also possible to rearrange the values in a vector by using indices. Look at the example code to see how indices work. 

*** =instructions
- Create the name vector.
- See how the indices work by executing the example lines.
- Create a new vector `girls` with values `Liisa` and `Anna` (in that order) from `names` by using indices.
- Create a new vector `boys` with values `Pekka` and `Matti` (in that order) from `names` by using indices.

*** =hint
- Note that space between the vector object and bracets produces an error.
- Remember that vectors are created with the function `c()`.
- Indices vectors `c(1,2)` and `c(2,1)` do not produce the same outcome. The order of the values is different.

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# Let's create a vector
names <- c('Matti', 'Pekka', 'Liisa', 'Anna')

# Acess the first value of the vector
names[1]

# Acess the 2. and 3. value of the vector
names[c(2, 3)]

# Rearranged
names[c(3, 2)]

# Separate the names vector into two vectors by using indices
girls <-
boys <-

```

*** =solution
```{r}
# Let's create a vector
names <- c('Matti', 'Pekka', 'Liisa', 'Anna')

# Acess the first value of the vector
names[1]

# Acess the 2. and 3. value of the vector
names[c(2, 3)]

# Rearranged
names[c(3, 2)]

# Separate the names vector into two vectors by using indices
girls <- names[c(3, 4)]
boys <- names[c(2, 1)]

```

*** =sct
```{r}

test_object("girls", incorrect_msg = "Did you create `girls`?")
test_object("boys", incorrect_msg = "Did you create `boys`? Are the values in the correct order?")

test_error()

# Final message the student will see upon completing the exercise
success_msg("Such indices. Wow. Much intelligent.")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:8e0ac16859

## Creating vectors

There are some convenient quick ways to create vectors with integer values (whole numbers) in R. The following two lines produce identical results

`1:5  
c(1,2,3,4,5)`

This is especially useful for creating vectors to use as index values.

There is also a quick and easy way to create a numeric vector of just zeroes with `numeric(length)`.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# df is available

# Create an integer vector
1:10

# Create an empty numeric vector
numeric(10)

# access var in df using indices
df$var[1:5]

```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:a26cc4ab51

## Looping (1)

In programming, often times there is the need to repeat an action multiple times. In statistical programming you might for example have a theory you wish to explore by simulation.

The structure of a for-loop is:

`for (counter in vector) {
  commands
}`

Inside the curly braces are regular commands such as function calls and the current value of the counter can be used there. The commands are repeated until all the values of the vector have been looped through. 



*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# loop through characters
for(letter in c("a","b","c")) {
  print(letter)
}

# loop through integers
for(i in 1:5) {
  print(i + 5)
}
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:b830d50ea8

## Looping (2)

A loop can be used to fill a vector with values. This is often useful because it is likely to want to store the results of looping.

One simple way to achieve this is to first create an empty vector using for example `numeric()`. The vector can then be filled with new values using indeces.


*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# Size of the experiment
N <- 10

# Create a numeric vector. Print out the values.
vec <- numeric(N)
vec

# fill the vector with new values
for(i in 1:N) {
  vec[i] <- i
}

# print out the new values
vec

```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:65d85c42aa

## The sampling distribution

Usually, whenever you have a set of data, that data is only a sample from a target population. Therefore all the statistics calculated from the data involve uncertainty. But how much uncertainty?  

The key to answering this question is the sampling distribution of the statistic.

Let's pretend for a moment that the data that we have actually is the target population. Let's see what happens if take samples from that population, calculate a statistic each time and store the results.


*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# df is available

# Create an empty vector
N <- 100
means <- numeric(N)

for(i in 1:N) {
  sample_data <- sample(df, 50)
  means[i] <- mean(sample_data$var)
 }
 
# Explore the sampling distribution
summary(means)
hist(means)

```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1

## The central limit theorem

Yleistet채채n edellisen teht채v채n huomio

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# df is available

#explore means
summary(means)
hist(means)

```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```
