---
title       : Hypothesis testing
description : UNDER CONSTRUCTION

--- type:NormalExercise lang:r xp:100 skills:1 key:ee77c10ab2

## Statistical hypothesis testing

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
learning2014 <-  read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/learning2014.txt", sep = "\t", header = TRUE)
```

*** =sample_code
```{r}
# sample code here

hist(learning2014$age, 35)
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:c033eeaa78

## Meet the tests! (1)

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:9b63d0cb6a

## Meet the tests! (2)

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:0b42a1089b

## Peculiar p-values

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:8dfc56d628

## (Null hypothesis:) which way to go?!

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

