---
title       : Exploring variation and dependence
description : Variation and dependence are at the heart of Statistics. In fact, without variation the discipline would cease to exist. Correlation is not causation, but why?

--- type:NormalExercise lang:r xp:100 skills:1 key:1d52143e10

## Guess the correlation

Correlation measures the strength of the *linear* relationship between two variables. It is important to note the word linear and try to think about what it means.

In an effort to make the concept more clear, we have build a game you can play.

*** =instructions
- Use `simulate()` to draw a scatter plot of the values of two simulated variables. The variables can have linear and nonlinear relationships. Try to guess what the correlation might be.
- Use `show_correlation()` to check how close you got.
- You can experiment as many times as you wish and the simulation will be different each time. Feel free to change `n`, the number of simulated values.

*** =hint
- When you are done experimenting, simply click 'Submit Answer' to move forward.


*** =pre_exercise_code
```{r}
# simulate and plot correlated data
simulate <- function(n = 60) {
  
  # degree (d) of the polynomial
  degree <- sample(1:3, 1)
  
  # x values
  x <- rnorm(n)
  
  # each x^d multiplied by it's coefficient
  # the coefficients are sampled from U(-1, 1)
  X <- sapply(1:degree, function(d) {
    b <- runif(1, min = -1, max = 1)
    b * x^d
  })
  
  # linear equation y = b_1 * x^1 + .. + b_d * x^d + e
  y <- rowSums(X) + rnorm(n)
  
  # scatterplot with regression line
  plot(x, y, col = "grey40", type = "p", pch= 20, xlab="", ylab="")
  
  return(list(x = x, y = y))
}

show_correlation <- function(df) {
  x <- df$x
  y <- df$y
  plot(x, y, type = "p", col = "grey40", pch= 20, xlab="", ylab="",
       sub = paste("correlation:", round(cor(x, y),2)))
  abline(lm(y~x), col ="red")
}

```

*** =sample_code
```{r}
# Functions simulate() and show_correlation() are available

# Simulate values and draw a scatterplot
data <- simulate(n = 60)

# Show the linear relationship and correlation
show_correlation(data)

```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}

test_error()
success_msg("Thanks for playing!")
```
