---
title       : Looking at the data
description : Data are everywhere and everything. That's why Statistics is also called Data Science. R offers great tools for looking at the data, behind the numbers.

--- type:NormalExercise lang:r xp:100 skills:1 key:9c7624db08

## Getting intimate with the data

In this section, you will learn how to get familiar with your data by visualizing and summarizing it.

These are always the first things to do after the data is collected and preprocessed (cleaned). These kinds of simple explorations are an important step and they create valuable insight into the data.


*** =instructions
- Take a look at the examples - you will learn more about all of them in the next exercises.
-  You can browse the different plots by clicking on "Next/Previous plot".
- Click 'Submit Answer' to move to the first exercise.

*** = hint
- Simply click 'Submit Answer' to move to the next exercise.


*** =pre_exercise_code
```{r}
  # load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
  # keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
  # recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
  # keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)

# bar plot of operating systems 
barplot(summary(students2014$kone), main = "Barplot of OS", xlab = "Linux, Mac, Win or none")

# histogram of heights
hist(students2014$pituus,  main ="Histogram of student heights", xlab = "height")
  
# horizontally placed box plot of ages
boxplot(students2014$ika, main = "Boxplot of student ages", xlab = "age", horizontal = TRUE)


```

*** =sample_code
```{r}
# students2014 is available

# Bar plot of operating systems 
barplot(summary(students2014$kone), main = "Barplot of OS", xlab = "Linux, Mac, Win or none")
    
# Histogram of student heights
hist(students2014$pituus,  main ="Histogram of student heights", xlab = "height")
  
# Box plot of student ages
boxplot(students2014$ika, main = "Boxplot of student ages", xlab = "age", horizontal = TRUE)


```

*** =solution
```{r}
# students2014 is available

# Bar plot of operating systems 
barplot(summary(students2014$kone), main = "Barplot of OS", xlab = "Linux, Mac, Win or none")
    
# Histogram of student heights
hist(students2014$pituus,  main ="Histogram of student heights", xlab = "height")
  
# Box plot of student ages
boxplot(students2014$ika, main = "Boxplot of student ages", xlab = "age", horizontal = TRUE)

```

*** =sct
```{r}
test_error()
success_msg("Ready? Let's go for it!")

```

--- type:NormalExercise lang:r xp:100 skills:1, 3  key:0e42548148
## Summary() statistics

Some functions in R are very generic. They can take different objects as their first argument and they seem to magically do exactly what you'd hope they do in all situations.

One such example is the `summary()` function. Depending on the type of data object it receives as it's first argument, `summary()` will produce different convenient results. 

*** =instructions
- Execute the code that creates a summary of `sukup`. What does it show you?
- Execute the code that creates a summary of `ika`. What does it show you?
- Create a summary of `pituus` in `students2014`.
- Create a summary of the object `students2014`.

*** =hint
- Use the `$` -sign to access `pituus` in `students2014`.

*** =pre_exercise_code
```{r}
# load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
# recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
# keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```


*** =sample_code
```{r}
# students2014 is available

# Create a summary of a factor
summary(students2014$sukup)

# Create a summary of a numeric
summary(students2014$ika)

# Create a summary of 'pituus' in 'students2014'


# Create a summary of the data.frame 'students2014'



```

*** =solution
```{r}

# Create a summary of a factor object
summary(students2014$sukup)

# Create a summary of a numeric object
summary(students2014$ika)

# Create a summary of 'pituus' in 'students2014'
summary(students2014$pituus)

# Create a summary of the data.frame 'students2014'
summary(students2014)



```

*** =sct
```{r}
test_output_contains("summary(students2014$pituus)", incorrect_msg = "Please call summary() on the numeric object students2014$pituus.")
test_output_contains("summary(students2014)", incorrect_msg="Please call summary() on the data.frame object `students2014`")

test_error()
success_msg("Well done!")
```


--- type:NormalExercise lang:r xp:100 skills:1 key:457280046c

## Bar plots of qualitative variables

Bar plots are useful for visualizing qualitative, discrete variables. A bar plot represents the counts of the unique values of a variable.

In this exercise you will draw a barplot of `kone`. For each student the value of `kone` is the operating system of the students computer. For more information see [the metafile](http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-meta.txt).

In base R graphics, the function `barplot()` can be used to draw a bar plot. The first argument (called `height`) should contain the counts you wish to visualize. The counts can be computed with `summary()` or more generally with `table()`.


*** =instructions
- Use `summary()` and `table()` on `kone` in the data frame `students2014`. Many of the values of `kone` are `<NA>`, not available. Why do you think this could be?
- Adjust the code `os_counts <- NULL`: Use `table()` and assign the counts of the different operating systems to `os_counts`.
- Use the function `barplot()` to draw a bar plot. Use the counts of the different operating systems as the first argument.

*** =hint
- Replace `NULL` with the example code where `table()` is used on `kone`.
- Give the object `os_counts` as the first argument to `barplot()`. 


*** =pre_exercise_code
```{r}
  # load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
  # keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
  # recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
  # keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```


*** =sample_code
```{r}
# students2014 is available

# Look at the counts of different levels of 'kone'.
summary(students2014$kone)
table(students2014$kone)

# Create object os_counts that stores the counts
os_counts <- NULL

# Call barplot() to visualize the counts.


  
```

*** =solution
```{r}
# students2014 is available

# The counts of different levels of 'kone'.
summary(students2014$kone)
table(students2014$kone)

# Create object os_counts that stores the counts
os_counts <- table(students2014$kone)

# Call barplot() to visualize the counts.
barplot(os_counts)



```

*** =sct
```{r}
test_object("os_counts", incorrect_msg = "Please use `table()` to create the object `os_counts`.")

test_function("barplot", args="height", not_called_msg = "Please call `barplot()` using the counts of the operating systems, created by `table()`.")

test_error()
success_msg("Excellent work!")
```

--- type:NormalExercise lang:r xp:150 skills:1 key:82217456fa

## Bar plots of continuous variables

When a variable is continuous by nature, it means that it can take on so many different values that all of it's observed values are likely to be unique. How to draw a bar plot in this case?

Age can be interpret as a continuous variable even though age is often measured quite discretely as rounded down years. Even then, usually there are too many unique values to visualize them clearly with a bar plot.

A solution is to first categorize the variable.


*** =instructions
- Use `table()` to look at the distribution of students age.
- Use `cut()` to create a factor object where age is cut into intervals. Use the vector `c(0, 16, 20, 24, 30, 100)` as break points.
- Use `table()` on the age categories to create object `age_counts` and print it.
- Visualize the contents of the table with a bar plot.


*** =hint
- Use the argument `breaks = c(0, 16, 20, 24, 30, 100)` inside `cut()` to create the factor object `age_cut`.
- Make sure to close the parenthesis on `cut()`.



*** =pre_exercise_code
```{r}
  # load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
  # keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
  # recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
  # keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```


*** =sample_code
```{r}
# students2014 is available

# Print out a table of the students ages
table(students2014$ika)

# Split age to intervals
age_cut <- cut(students2014$ika, breaks = NULL)

# Create a table of the students age categories and print it.
age_counts <- NULL


# Draw a barplot() of the counts of the age categories. 



```

*** =solution
```{r}
# students2014 is available

# Print out a table of the students ages
table(students2014$ika)

# Split age to intervals
age_cut <- cut(students2014$ika, breaks = c(0, 16, 20, 24, 30, 100))

# Create a table of the students age categories and print it.
age_counts <- table(age_cut)
age_counts

# Draw a barplot() of the age categories
barplot(age_counts)


```

*** =sct
```{r}

test_object("age_cut", incorrect_msg = "Please use `breaks = c(0, 16, 20, 24, 30, 100)` to create the object age_cut.")

test_object("age_counts",incorrect_msg = "Please create the object `age_counts` using `table()`.")

test_output_contains("age_counts", incorrect_msg = "Please print out the contents of `age_counts`.")

test_function("barplot", args = "height", not_called_msg = "Please use `barplot()` to visualize the counts of the age categories.")

test_error()
success_msg("Nicely done. That was awsome!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:86cee4bd71

## Histograms

Splitting a continuous variable into intervals of equal width and then drawing a bar plot is so common that the procedure has it's own name: the histogram. 

In R, basic histograms can be drawn with the `hist()` function. The first argument of `hist()` should be a numeric vector of data values. The argument `breaks` can be used to approximately control the number of intervals.

The plotting functions in R also have many additional arguments which allow for more control over the plot. One simple argument is `col`, which can be the number(s) or the name(s) of the desired color(s). Some possible color names can be found <a target="_blank" href="http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf">here</a>.

*** =instructions
- Adjust the sample code and draw the histogram again by setting the argument `breaks = 20`.
- Use the `col` argument to give the histogram a nice color of your choice. Use quotations marks to define the color name.
- Set the argument `xlab = "height"`.
- Set the argument `main = "My histogram"`


*** =hint
- Remember to seperate the arguments of a function with a comma.
- Use quotation marks for the color names and the main and xlab titles.


*** =pre_exercise_code
```{r}
# load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
# recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
# keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)

# Draw a histogram of student heights
hist(students2014$pituus)

```

*** =sample_code
```{r}
# students2014 is available

# Draw a histogram of student heights
hist(students2014$pituus)



```

*** =solution
```{r}
# students2014 is available

# Draw a histogram of student heights
hist(students2014$pituus, breaks = 20, col = "blue", xlab = "height", main = "My histogram")


```

*** =sct
```{r}
# submission correctness tests

test_function("hist", args=c("x","breaks","main","xlab"))

test_error()
success_msg("Very good, young padawan, but you still have much to learn.")

```



--- type:NormalExercise lang:r xp:100 skills:1 key:c3d62c54f5

## Box plots

One very good way to visualize the summary statistics of a numerical variable is by drawing box plots. A box plot visualizes the 25th, 50th and 75th percentiles (the box), the typical range (the whiskers) and the outliers (the dots) of a variable. 

The whiskers extending from the box can be computed by several techniques. The default in R is to extend them to reach to a data point that is no more than 1.5*IQR away from the box, where IQR is the inter quartile range defined as  

`IQR = 75th percentile - 25th percentile`  

Values outside the whiskers can be considered as outliers, unusually distant observations. For more information on IQR, see <a target="_blank" href ="https://en.wikipedia.org/wiki/Interquartile_range"> wikipedia</a>

*** =instructions
- Create a summary of `students2014$ika`.
- Use `boxplot()` on `students2014$ika`.
- Adjust the plot: set the argument `horizontal = TRUE`.
- Adjust the plot: set the argument `xlab = "age"`.
- Give your plot the main title "Box plot of student ages" with the argument `main`.

*** =hint
- Give `students2014$ika` as the first argument of `boxplot()`.
- Remember to separate function arguments with a comma.
- Use quotation marks when adding text.

*** =pre_exercise_code
```{r}
# load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
# recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
# keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```


*** =sample_code
```{r}
# students2014 is available

# Create a summary of student ages
summary(students2014$ika)

# Visualize the distribution of student ages with a boxplot()




```

*** =solution
```{r}
# students2014 is available

# Create a summary of student ages
summary(students2014$ika)

# Visualize the distribution of student ages with a boxplot()
boxplot(students2014$ika, horizontal = T, xlab = "age", main = "Box plot of student ages")



```

*** =sct
```{r}

test_function("boxplot", args=c("x","horizontal","xlab","main"))

test_error()
success_msg("That was nice work!")

```

