---
title       : R data and variable types
description : Quick overview to R data and variable types.

--- type:NormalExercise lang:r xp:100 skills:1 key:e4360dcfec
## Data types in R (1)

There are different data types in R. The most commonly used are
- Numbers and decimal values are called numeric
- Text or character values are called characters
- Boolean values (`TRUE` or `FALSE`) are called logical

*** =instructions
- Create character object `my_initials` by assigning your initials in it.
- Assign the number 10.5 in `my_numeric_object`. Note that the decimal separator in R is dot instead of comma. 
- Create the object `my_boolean` by executing the line. Note that when creating boolean values, no quotation marks are needed.
- Change the value of `my_boolean` to `TRUE`
- str()?

*** =hint
- When creating a character object, remember to use quotation marks.
- Dot is the decimal mark in R:  `2.5` is decimal number but `2,5` means a different thing!
- Remember that `Ctrl + Enter` executes a single row.

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# Create objects with different data types
my_initials<-
my_numeric_object<-
my_boolean<-FALSE

# Change the value of my_boolean to TRUE


```

*** =solution
```{r}
# Create objects with different data types
my_initials<-"E.K."
my_numeric_object<-10.5
my_boolean<-FALSE

# Change the value of my_boolean
my_boolean<-TRUE
```

*** =sct
```{r}
test_student_typed("my_boolean<-TRUE", not_typed_msg="Something is not quite right with my_boolean")
test_object("my_initials")
test_student_typed("my_numeric_object<-10.5", not_typed_msg="Object my_numeric is incorrect")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:9ee1ba578e
## Data types in R (2)

- Factors and `cut()`

*** =instructions
- 

*** =hint
- 

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
```

*** =solution
```{r}
```

*** =sct
```{r}
# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:8718babe2e
## Data frames

The most common object to operate on is the `data frame` object, which stores data in tabular form.

In R, the basic idea is to use *functions* to operate on objects. Here you will get to apply some functions that are helpful for inspecting data: `head()` and `str()`.

*** =instructions
- Execute the function `head()` and see the results
- Inspect the data types in the `JYT` data with the function `str()`

*** =hint
- The data is stored in object `JYT`

*** =pre_exercise_code
```{r}
source("functions.R")
JYT<-getdata()
```

*** =sample_code
```{r}
# Let's view the first 6 students from the course participants data
head(JYT)

# View the data types in the JYT data


```

*** =solution
```{r}
# Let's view the first 6 students from the course participants data
head(JYT)

# View the data types in the JYT data
str(JYT)
```

*** =sct
```{r}
test_function("head")
test_function("str")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:7cc56d2158
## Histogram


*** =instructions
- 

*** =hint
- 

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
```

*** =solution
```{r}
```

*** =sct
```{r}
# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:3c0486635f
## Barplot


*** =instructions
- 

*** =hint
- 

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
```

*** =solution
```{r}
```

*** =sct
```{r}
# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


