---
  title       : Estimation
description : UNDER CONSTRUCTION

--- type:NormalExercise lang:r xp:100 skills:1
## Exploring estimates with R

In this chapter we will focus on point and interval estimation. Interval estimation is based on the important concept of the *sampling distribution*, which is a somewhat theoretical concept and therefore takes a while to get used to.

You can think of the sampling distribution as a kind of mind game: When we have a random sample of data, we can calculate statistics such as the mean of that sample. But what if we could collect the random sample again and again and again and calculate the mean for each of the samples. What would then be the distribution of these calculated means? This theoretical distribution is called the sampling distribution.

It turns out that for a lot of important point estimates, the sampling distribution can be approximated with the normal distribution with certain known parameters! This is due to a fascinating phenomenom called the *central limit theorem*.

To explore these concepts, we will start this chapter with some technical tools we will need for constructing *loops* and *simulations*. 

*** =instructions
- Here in DataCamp, when a single expression continues over multiple rows, the whole chunk of code needs to be executed at once. This can be done by first painting (selecting) all the lines in the expression with a mouse and then pressing `Ctrl + Enter` as normal.
- Execute the while for-loop expression ending to the line where the curly bracket closes
- Click submit answer to move on to the next exercise

*** =hint
- If you cannot paint (select) the expression for some reason, you can just click submit answer instead.

*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# Execute me
for(i in 1:10) {
  print("I am a loop")
}




```

*** =solution
```{r}
# Execute me
for(i in 1:10) {
  print("I am a loop")
}




```

*** =sct
```{r}
test_error()
success_msg("Good work!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:23abc9f934

## Indices and brackets

Vectors in R store multiple values of the same data type. The values in vectors have indices: the first value in a vector has a index `1`, the second value `2` and so on. 

You can set or get a single value from a vector by using indices and brackets `[ ]`. Using an index number inside the bracets gives access to a single value from the vector. Using a vector of indices gives access to multiple values (another vector). Therefore it is also possible to rearrange the values in a vector by using indices. Look at the example code to see how indices work. 

*** =instructions
- Create the name vector.
- See how the indices work by executing the example lines.
- Use brackets and indices on `names` to create a new vector `girls` with values "Liisa" and "Anna" (in that order).
- Use brackets and indices on `names` to create a new vector `boys` with values "Pekka" and "Heikki" (in that order)

*** =hint
- Note that space between the vector object and bracets produces an error.
- Index vectors `c(1,2)` and `c(2,1)` do not produce the same outcome. The order of the values is different.

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# Let's create a vector
names <- c("Matti", "Pekka", "Liisa", "Anna")

# Acess the first value of the vector
names[1]

# Change the first value of the vector
names[1] <- "Heikki"

# Acess the 1. and 3. value of the vector
names[c(1, 3)]

# Use indices and brackets to separate the names vector into two vectors of the specified ordering
girls <-
boys <-


  ```

*** =solution
```{r}
# Let's create a vector
names <- c("Matti", "Pekka", "Liisa", "Anna")

# Acess the first value of the vector
names[1]

# Change the first value of the vector
names[1] <- "Heikki"

# Acess the 1. and 3. value of the vector
names[c(1, 3)]

# Use indices and brackets to separate the names vector into two vectors of the specified ordering
girls <- names[c(3, 4)]
boys <- names[c(2, 1)]

```

*** =sct
```{r}

test_object("girls", incorrect_msg = "Did you create `girls`?")
test_object("boys", incorrect_msg = "Did you create `boys`? Are the values in the correct order?")

test_error()

# Final message the student will see upon completing the exercise
success_msg("Such indices. Wow. Much intelligent.")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:8e0ac16859

## Easy vectors

Sometimes you need a very long index vector. In that case it would be a lot of work to create the vector by combining integer values with `c()`. Luckily, there are convenient ways to create long vectors in R.

For index values, a specially useful one is the method `start:end`, which creates an integer vector with all the values from start to end. The following two lines hence produce identical results

```
1:5  
c(1,2,3,4,5)
```


*** =instructions
- Study and execute the example codes
- Use `:` to create and print out an integer vector with the values 1, 2, ..., 10
- Print out every second value of the attitude vector, starting from the 2. value until the 20th value. These values correspond to the even numbered indeces of the vector: 2, 4, .. , 20

*** =hint
- First you will need an index vector with values 2, 4, .. , 20. The example shows how to create such a vector
- You can then use the index vector together with brackets (`[ ]`) to complete the task


*** =pre_exercise_code
```{r}
learning2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/learning2014.txt", sep = "\t", header = TRUE)
attitude <- learning2014$attitude
names(attitude) <- 1:length(attitude)
```

*** =sample_code
```{r}
# attitude is available

# Create and print an integer vector with values 1,2,..,5
1:5

# Create an integer vector with even values 2, 4, .. , 10
(1:5)*2

# Access the values 5 - 10 of attitude
attitude[5:10]

# Create and print an integer vector with values 1,2,..,10


# Access every second value of attitude from 2. to the 20th index


```

*** =solution
```{r}
# attitude is available

# Create an integer vector
1:5

# Create an integer vector with even values
(1:5)*2

# Access the values 5 - 10 of attitude
attitude[5:10]

# Create and print an integer vector with values 1,2,..,10
1:10

# Access every second value of attitude from 2. to the 20th index
attitude[(1:10)*2]


```

*** =sct
```{r}
# submission correctness tests

test_student_typed("1:10", not_typed_msg = "Did you use `:` to create and print out the specified integer vector?")
test_output_contains("attitude[(1:10)*2]")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Great work!")

```

--- type:NormalExercise lang:r xp:100 skills:1  key:78e301b70a

## Point estimation

Let us now get back to statistical analysis and estimation. When given a sample of data from a population, the goal of statistical analysis is to use that sample to draw conclusions about the population. In a simple case we are interested in a single variable (such as students points in an exam) and want to estimate the *expected value* of that variable. The expected value is the average value in the target population - the population mean. 

We can't directly calculate the population average because we only have a sample, so we have to **estimate** it. Obtaining a point estimate of a population parameter is rather easy: just use the corresponding sample statistic.

population parameter                    | estimate
--------------------------------------- | --------
expected value $\mu$                    | sample mean $\bar{x}$
population standard deviation $\sigma$  | sample standard deviation $s$

In statistics, estimates are often denoted with a hat. So for example $\hat{\mu} = \bar{x}$.

*** =instructions
- Estimate the expected value of `points`. Remember that R has convenient functions for these kind of calculations. If you can't remember them, use your favourite search engine or take a hint.
- Estimate the population standard deviation of `points`
- Combine the estimates to the estimates vector (replace `NA`). Notice how `c()` can be used to give names to the values.

*** =hint
- The table above shows the relationships between the sample statistics and the population parameters.
- Use the table to figure out which operation you could use to produce the estimate.



*** =pre_exercise_code
```{r}
learning2014 <-  read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/learning2014.txt", sep = "\t", header = TRUE)
```

*** =sample_code
```{r}
# learning2014 is available

# create object points using the learning2014 data.frame
points <- learning2014$points

# estimate the expected value of points
mu_hat <- 
  
# estimate the population standard deviation of points
sigma_hat <- sd(points)

# Combine the estimates to a named vector and print out the rounded values
estimates <- c("mu_hat" = NA, "sigma_hat" = NA)
round(estimates, 2)



```

*** =solution
```{r}
# learning2014 is available

# create object points using the learning2014 data.frame
points <- learning2014$points

# estimate the expected value of points
mu_hat <- mean(points)

# estimate the population standard deviation of points
sigma_hat <- sd(points)

# Print out the estimated values
estimates <- c("mu_hat" = mu_hat, "sigma_hat" = sigma_hat)
round(estimates, 2)


```

*** =sct
```{r}
# submission correctness tests

test_function("mean", args=c("x"))
test_object("mu_hat", incorrect_msg = "Please create the object mu_hat. Use the correct function on the points vector")

test_output_contains("round(estimates, 2)", incorrect_msg = "Please insert the estimate objects to the estimates vector and do ont remove the row with `round()`")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Very nice! You get full points for point estimation.")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:1df061a41a

## Interval estimation

Since we are using a random sample, clearly there is uncertainty involved in our point estimate. But how much uncertainty? This is where we need statistical theory. One way to approach this question is to try to come up with an interval around the the point estimate which we could be confident that the interval contains the true parameter value a high percentage of the time. 

It turns out that the following formula produces an interval which contains the true population mean a specified high percentage of the time ($100 \cdot (1-\alpha)$%) 

$$\text{mean} +- z(\alpha/2) \cdot \frac{s}{\sqrt{n}}$$
  
where $s$ is the sample standard deviation, $n$ is the sample size and $z(\frac{\alpha}{2})$ is a critical value from the $N(0,1)$ distribution. $\frac{s}{\sqrt{n}}$ is called the *standard error* of the estimate. 

I know what you're thinking. It's confusing: where did the normal distribution come from, what's with all that $z(\frac{\alpha}{2})$ stuff and what critical value? All the answers lie in the **sampling distribution** which we will soon get to.


*** =instructions
- Compute the standard error of the sample mean
- Compute the critical value z using a 99% confidence level (alpha = 0.01)
- Round the values to 1 digit and print out the interval estimate

*** =hint
- The formula for the standard error is s/sqrt(n)
- The function qnorm() computes quantile values from the N(0,1) distribution. Use it as instructed in the code comments.


*** =pre_exercise_code
```{r}
learning2014 <-  read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/learning2014.txt", sep = "\t", header = TRUE)

points <- learning2014$points

```

*** =sample_code
```{r}
# points is available.

# Number of observations
n <- length(points)

# Sample mean and standard deviation of points
mu_hat <- mean(points)
s <- sd(points)

# Compute the standard error of the sample mean. sqrt() computes the square root.
error <- 

# Compute the critical value z
z <- qnorm(0.01/2, lower.tail = F) 

# Compute the confidence interval
lower_ci <- mu_hat - NA
upper_ci <- mu_hat + NA

# Combine, round and print out the results
interval_estimate <- c("estimate" = mu_hat, "lower99%" = lower_ci, "upper99%" = upper_ci)



```

*** =solution
```{r}
# points is available

# number of observations
n <- length(points)

# estimate the expected value and standard deviation of points
mu_hat <- mean(points)
s <- sd(points)

# Compute the standard error of the sample mean. sqrt() computes the square root.
error <- s/sqrt(n)

# Use qnorm(1 - alpha/2) to get the value of z with the specified confidence level
z <- qnorm(1 - 0.01/2)

# compute the confidence interval and print out the results
lower_ci <- mu_hat - z*error
upper_ci <- mu_hat + z*error

# Combine, round and print out the results
interval_estimate <- c("estimate" = mu_hat, "lower_ci" = lower_ci, "upper_ci" = upper_ci)
round(interval_estimate)


```

*** =sct
```{r}
# submission correctness tests
test_object("error")
test_object("lower_ci")
test_object("upper_ci")
test_output_contains("round(interval_estimate)")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("You are awsome with 99% confidence!")


```


--- type:NormalExercise lang:r xp:100 skills:1 key:a26cc4ab51

## Looping

In programming, often there is the need to repeat an action multiple times. In statistical programming you might for example have a theory you wish to explore by simulation.

The **for-loop** is perhaps the most common loop. The structure of a for-loop is:

```
for (counter in vector) {
  commands
  more commands
}
```

The for-loop iterates trough the values of a vector by changing the value of the counter one value at a time. The counter first takes on the first and value of th vector, then the second and so on. Inside the curly braces is the *body* of the loop, which can contain regular commands such as function calls. The current value of the counter can be used there.

The commands are repeated until the counter has taken on all the values of the vector.


*** =instructions
- Execute the example codes. Remember that here in DataCamp, executing a command over multiple lines is done by selecting all the lines with a mouse first and then hitting `Ctrl+Enter` normally. Alternatively you could also click submit to execute the whole script.R
- In an honor of the electronic duo Daft Punk and their 2001 Album "Discovery", construct a for-loop that prints out "One more time!" 27 times.
- Hints: (1) Remember that `1:n` creates an integer vector of length n. (2) Inside the loop body (curly braces) you have to use the `print()` function to print.

*** =hint
- It does not matter what name you give to your counter. You can for example use `i` as is done in the second example.
- In this exercise you don't need to use the values of your counter.
- Remember to use quotation marks to print characters


*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Loop through and print the characters a, b, c, d
for(letter in c("a","b","c")) {
  print(letter)
}

# Loop through integers 1, 2, 3, 4, 5
for(i in 1:5) {
  print(i + 5)
}

# Write a for-loop that prints out "One more time!" 27 times





```

*** =solution
```{r}
# loop through characters a,b,c,d
for(letter in c("a","b","c")) {
  print(letter)
}

# loop through integers 1, 2, 3, 4, 5
for(i in 1:5) {
  print(i + 5)
}

# Write a for-loop that prints out "One more time!" 27 times
for(i in 1:27) {
  print("One more time!")
}


```

*** =sct
```{r}
# submission correctness tests

test_output_contains("'One more time!'", times = 27, incorrect_msg = "Did you write a for-loop that prints out 'One more time!' 27 times?")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Excellent work! Repetition is the key to success :)")

```



--- type:NormalExercise lang:r xp:100 skills:1 key:65d85c42aa

## The sampling distribution

Remember that you can think of the sampling distribution as a kind of mind game:  If we could collect our random sample again and again and again and calculate a statistic for each of the samples the resulting distribution would give us the sampling distribution of that statistic.

In reality, we can't ofcourse produce this distribution, or can we? 

Let's pretend for a moment that the data we have actually *is* the target population we are sampling from. Let's see what happens if we repeatedly take samples from our data, and calculate a statistic each time, storing the results. This satisfies the idea of the sampling distribution, meaning that we can actually simulate that distribution! 

This method (called [bootstrapping](https://en.wikipedia.org/wiki/Bootstrapping_(statistics))) is widely used to explore different sampling distribution.


*** =instructions
- Print out the empty means vector
- Execute the for-loop. What does it do?
- Print out the means vector
- Use `summary()` on means
- Draw a histogram of the sampling disribution of the means. See the help page of hist and make it a density histogram.


*** =hint
- `hist()` draws a histogram and freq = F can be used to draw a density histogram


*** =pre_exercise_code
```{r}
learning2014 <-  read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/learning2014.txt", sep = "\t", header = TRUE)
points <- learning2014$points

```

*** =sample_code
```{r}
# points is available. Do not remove the line below.
set.seed(777)

# Create an empty vector of length N
N <- 100
means <- numeric(N)

# Print out the empty vector


# Repeat N times: (1) draw a random sample, (2) compute the mean and (3) store it in the means vector
for(i in 1:N) {
  points_sample <- sample(points, size = 50, replace = F)
  means[i] <- mean(points_sample)
}

# Explore the sampling distribution of the mean




```

*** =solution
```{r}
# points is available. Do not remove the line below.
set.seed(777)

# Create an empty vector of length N
N <- 100
means <- numeric(N)

# Print out the empty vector
means

# Repeat N times: (1) draw a random sample, (2) compute the mean and (3) store it in the means vector
for(i in 1:N) {
  points_sample <- sample(points, size = 50, replace = F)
  means[i] <- mean(points_sample)
}

# Explore the sampling distribution
means
summary(means)
hist(means, freq = F)

```

*** =sct
```{r}
# submission correctness tests

test_output_contains("means")
test_function("summary", args=c("object"))
test_function("hist", args=c("x","freq"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:38bf66cd9c

## The central limit theorem

Yleistet채채n edellisen teht채v채n huomio

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
learning2014 <-  read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/learning2014.txt", sep = "\t", header = TRUE)
points <- learning2014$points
set.seed(777)

# Create an empty vector of length N
N <- 100
means <- numeric(N)

# Repeat N times: (1) draw a random sample, (2) compute the mean and (3) store it in the means vector
for(i in 1:N) {
  points_sample <- sample(points, size = 50, replace = F)
  means[i] <- mean(points_sample)
}

# Explore the sampling distribution
hist(means, freq = F)
```

*** =sample_code
```{r}
# df is available

N <- 100

# Repeat N times: (1) draw a random sample, (2) compute the mean and (3) store it in the means vector
for(i in 1:N) {
  points_sample <- sample(points, size = 50, replace = F)
  means[i] <- mean(points_sample)
}



```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```
