---
title       : R and statistics
description : Basics of R, the amazing statistical programming language.

--- type:NormalExercise lang:r xp:50 skills:1 key:168121d0f0

## What is R?

R is a programming language designed for statistical analysis. R is completely open source and all analysis done with R are reproduciple and easy to share. R is very flexible and powerful and always includes implementations of the latest research. 

Because of these reasons, R is a very popular language among the scientific community. R is however also widely [used]("http://blog.revolutionanalytics.com/2014/05/companies-using-r-in-2014.html") for data analysis and visualisations by companies such as Google, Facebook, Twitter and Reaktor. R is perfect for data science.

Researchers and experts often use programming to visualize and explore data and implement new methods. The purpose of this course is to expose you to the basics of programming and statistical analysis with R.


*** =pre_exercise_code
```{r}
# pre exercise code here
data_url <- "https://raw.githubusercontent.com/TuomoNieminen/Helsinki-Social-Statistics/master/data/JYTOPKYS-data.txt"
JYT <- read.table(data_url, sep="\t", header=TRUE)
with(JYT, 
     plot(kenka, pituus, col = sukup))
```

*** =sample_code
```{r}
# This is a very simple showing some of the power of R.

# Load a dataset of 2014 JYT course students from the web and store the data by creating an object named 'JYT'.
data_url <- "https://raw.githubusercontent.com/TuomoNieminen/Helsinki-Social-Statistics/master/data/JYTOPKYS-data.txt"
JYT <- read.table(data_url, sep="\t", header=TRUE)

# Plot shoesize versus hight using functions with() and plot(). Indicate sex by color using the col argument of plot().
with(JYT, 
     plot(x = kenka, y = pituus, col = sukup))

```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

success_msg("Ok, let's get started! Move on to the first exercise.")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:35ad6c43ec
## Basic tools

On your right you see the R editor area - the script - and below that the R console. The editor area is just a simple text editor where you write R code.

The way to use these two is to first write code to the editor area and then to tell R that you want to execute a line of code by pressing `Ctrl + Enter`. This executes the line where your cursor currently is. Input and output will then appear in the console.

Here at DataCamp you can also execute everything in the editor by pressing 'Submit answer', after which DataCamp will do some magic to check if you have correctly completed the exercise.

Try it!
  
*** =instructions
- Type "Hello world!" in the editor area. Use quotation marks.
- Press `Ctrl + Enter` to execute your "Hello world!" - code
- Click 'Submit answer' when done.

*** =hint
- Remember to use quotation marks.


*** =pre_exercise_code
```{r}
# no pec
```


*** =sample_code
```{r}
# This is the R editor!

# A hashtag at the beginning of the line defines the line as a comment

# Write code below



# Below is the R console, where you will see output
```

*** =solution
```{r}
# This is the R editor!

# A hashtag at the beginning of the line defines the line as a comment

# write your code below
"Hello world!"


# Below is the R console, where you will see output
```

*** =sct
```{r}

test_student_typed("Hello world!", not_typed_msg="Please type 'Hello world!' in the editor.")

test_error()

# Final message the student will see upon completing the exercise
success_msg("Great work! You have now executed your first line of code :)")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:b0a54af3ff
## Arithmetics with R

Let's continue with something simple. R can do amazing things like scrape websites and draw beautiful graphics but it can also do basic calculations. Consider the following arithmetic operators:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`

The `^` operator raises the number to its left to the power of the number to its right: for example `3^2` is 9. With this knowledge, follow the instructions below to complete the exercise.

*** =instructions
- Type `2^5` in the editor to calculate 2 to the power 5.
- Use `Ctrl + Enter` to run the line and/or click 'Submit answer' and have a look at the R output in the console.
- Note how the `#` symbol is used to add comments.


*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# An addition
5 + 5 

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation



```

*** =solution
```{r}
# An addition
5 + 5

# A subtraction
5 - 5 

# A multiplication
3 * 5

 # A division
(5 + 5) / 2 

# Exponentiation
2 ^ 5

```

*** =sct
```{r}

test_output_contains("2^5", incorrect_msg = "The exponentiation example is not correct. Please write `2 ^ 5` on a new line.")

success_msg("Great work! Head over to the next exercise.")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:856fab934f
## Objects

Here's where things start to get interesting. In R you can create and operate on things called *objects*. 
An object is something that can store information such as numerical values and names. Once an object is created, it will be stored in memory and the information it contains will be available to you later.

Objects are created using the assign operator: `<-`. The value of an object can be printed by typing it's name.


*** =instructions
- Execute the lines that create and operate on `my_hello_object` and `my_numeric_object`.
- Assign your name or nickname to `my_name`. Use quotation marks.
- Assign a positive message to yourself to `my_message`. Use quotation marks.
- Click 'Submit answer'.

*** =hint
- Remember that `Ctrl + Enter` executes a single row.

*** =pre_exercise_code
```{r}
# no pec
```


*** =sample_code
```{r}

# Create an object
my_hello_object <- "Hi there!"

# Print the contents of the object
my_hello_object

# Create another object
my_numeric_object <- 5 + 5

# Do further calculations with the object
my_numeric_object / 5

# Override the value of an object by assigning a new value to it
my_hello_object <- my_numeric_object
my_hello_object

# Create objects my_name and my_message here. Use quotation marks.
my_name <- 
my_message <- 

```

*** =solution
```{r}

# Create an object
my_hello_object <- "Hi there!"

# Print the contents of the object
my_hello_object

# Create another object
my_numeric_object <- 5 + 5

# Do further calculations with the object
my_numeric_object / 5

# Override the value of an object by assigning a new value to it
my_hello_object <- my_numeric_object
my_hello_object

# Create objects my_name and my_message here. Use quotation marks.
my_name <- "Tuomo"
my_message <- "You are awsome!"

```

*** =sct
```{r}

test_object("my_name", eval = F, incorrect_msg = "Did you assign your name to `my_name`?")

test_object("my_message", eval = F, incorrect_msg = "Did you assign a message to `my_message`?")

test_error()

success_msg(paste0("Great job ", my_name, ". ", my_message, "!"))

```

--- type:NormalExercise lang:r xp:100 skills:1 key:b81cdce659
## Functions and arguments

In R you operate on objects using *functions*. A function is a special kind of object that can be *called* by typing it's name and giving it the necessary *arguments* inside parenthesis. This is how you're actually going to do all the magic.

During this course you will operate on ready-made functions, but for clarity we will look at the definition of a simple function.


*** =instructions
- Study the definition of the add_k() function.
- Execute the function calls that are defined.
- Create object 'ans' by directly assigning the output of add_k() to it. Call add_k() with value = 10 and k = 5.

*** =hint
- The answer is of the form: `ans <- add_k(arguments here)`

*** =pre_exercise_code
```{r}
add_k <- function(value, k = 5) {
  newvalue <- value + k
  newvalue
  }
```


*** =sample_code
```{r}

# Below is a definition of a function: add_k(). It is a simple function with two arguments 'value' and 'k', one of which has a default value.
add_k <- function(value, k = 5) {
  newvalue <- value + k
  newvalue
  }

# Call the add_k() function twice.
add_k(value = 5)
add_k(value = 5, k = -4)

# If you give the arguments in the right order, you don't have to state which is which.
add_k(5, -4) # is equivalent to
add_k(k = -4, value = 5)

# Call add_k() to create the object 'ans'.
ans <- 

```

*** =solution
```{r}

# Call the add_k() function twice.
add_k(value = 5)
add_k(value = 5, k = -4)

# If you give the arguments in the right order, you don't have to state which is which.
add_k(5, -4) # is equivalent to
add_k(k = -4, value = 5)

# Call add_k() to create the object 'ans'.
ans <- add_k(value = 10)

```

*** =sct
```{r}

# tests

test_function("add_k",args=c("value","k"))
test_object("ans")

test_error()

success_msg("Great job! Now let's use some real functions!")

```

--- type:MultipleChoiceExercise lang:r xp:100 skills:1 key:8d2100f40a

## Meet the data

In future exercises, we will be doing some data science using a dataset that was collected from the students of this course in 2014. 

The students filled out [this short questionnaire]("https://github.com/TuomoNieminen/Helsinki-Social-Statistics/blob/master/data/ASSIST-2014-FI-0-20140903.pdf"). Please go ahead and read both pages of the pdf in the above link and then come back and answer the question below.

How many numbered sections does the questionnaire have?

*** =instructions
- two
- four
- seven
- fourteen

*** =hint
- if the link does not work, the url is:  https://github.com/TuomoNieminen/Helsinki-Social-Statistics/blob/master/data/ASSIST-2014-FI-0-20140903.pdf
- What is the number of the last section with background questions?


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sct
```{r}
# submission correctness tests
msg1 <- msg3 <- msg4 <- "Try again"
msg2 <- "Yes, correct!"
test_mc(correct = 2, feedback_msgs = c(msg1, msg2, msg3, msg4))

# Final message the student will see upon completing the exercise
success_msg("Great work! Next time we'll explore the data :)")

```
