---
title: 'Exploring variation and dependence'
description: 'Variation and dependence are at the heart of Statistics. In fact, without variation the discipline would cease to exist. Correlation is not causation, but why?'
---

## Be aware of varying variables

```yaml
type: NormalExercise
key: ea11f25c3e
lang: r
xp: 100
skills: 1
```

In the previous chapter you learned how to explore the variables in your data by creating summaries and plots out of data objects in R. You already know quite a bit about how to exlore variability.

In this chapter we'll expand on these tools and we'll also look at how to explore the *relationships* between pairs of variables with scatter plots and correlation. We'll also learn about standard deviation, perhaps the most important measure of variability in statistics.

`@instructions`
- Draw a scatter plot of students shoesizes and heights with `plot()`. Does a linear relationship between the variables look plausible?
- Compute the correlation between shoesizes and heights. What is your interpretation?
- Click Submit Answer to move to the next exercise.

`@hint`
- Simply execute the example codes and click Submit Answer.

`@pre_exercise_code`
```{r}
# load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
# recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
# keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)

```

`@sample_code`
```{r}
# students 2014 is available

# Create objects shoesize and height
shoesize <- students2014$kenka
height <- students2014$pituus

# Draw a scatter plot of shoesize and height
plot(shoesize, height)

# Compute the correlation of shoesize and height
cor(shoesize, height)


```

`@solution`
```{r}
# students 2014 is available

# Create objects shoesize and height
shoesize <- students2014$kenka
height <- students2014$pituus

# Draw a scatter plot of shoesize and height
plot(shoesize, height)

# Compute the correlation of shoesize and height
cor(shoesize, height)

```

`@sct`
```{r}

test_error()
success_msg("Good job!")

```

---

## Help() !

```yaml
type: NormalExercise
key: ab9cf651a8
lang: r
xp: 100
skills: 1
```

There are many functions in R. It is often hard to remember all the argument names and details of the functions. Luckily, every single function has a help page. Using the help pages is part of using R.

The help page of a function called *func* can be opened by two methods: `help(func)` or `?func`. The help page is a pop-up window, which you might have to specifically allow to pop up.

`@instructions`
- Open the help page of `plot` by writing `?plot` directly to the console and pressing enter.
- Draw a scatter plot of student shoesize versus height.
- Compute the correlation of student shoesize and height using `cor()`.
- Use the information on the help page of `plot()` and make the following adjustments to the plot:
- Add this (exact) main title: "Scatter plot of student shoesize versus height".
- Add this subtitle: "Correlation: 0.xx". Replace 0.xx with the actual correlation, rounded to two decimal places.

`@hint`
- Please note that the tests are sensitive to capital letters and missing spaces etc.
- See the previous exersize for tips on drawing a scatterplot and computing the correlation.

`@pre_exercise_code`
```{r}
# load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
# recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
# keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)

```

`@sample_code`
```{r}
# Create objects shoesize and height
shoesize <- students2014$kenka
height <- students2014$pituus

# Draw a scatterplot of shoesize and height


# Compute the correlation of shoesize and height


```

`@solution`
```{r}
# Create objects shoesize and height
shoesize <- students2014$kenka
height <- students2014$pituus

# Draw a scatterplot of shoesize and height
plot(shoesize, height, main = "Scatter plot of student shoesize versus height", sub = "Correlation: 0.85")

# Compute the correlation of shoesize and height
cor(shoesize, height)


```

`@sct`
```{r}

test_function("plot",args = c("x","y","main","sub"))

test_error()
success_msg("Very good work!")
```

---

## Standard deviation

```yaml
type: NormalExercise
key: 0083627e50
lang: r
xp: 100
skills: 1
```

*Standard deviation* is perhaps the most important measure of variability in statistics. It is defined as the average distance of a set of data points from their mean. So, the closer the values of a variable are to their mean on average, the smaller the standard deviation.

Mathematically, the sample standard deviation ($sd$) is computed using  

$sd = \sqrt[2]{\frac{ \Sigma (x - mean)^2}{n - 1} }$  

which might look a bit complicated but really is just the average distance from the mean. In R, standard deviation can be computed using the function `sd()`.

`@instructions`
- (1) Use `simulate()` to visualize a random sample of data from a population where the average distance from the mean is 1 (`sd = 1`). Compute the sample standard deviation with `sd()`.
- (2) Create object `x2` similarily by setting `sd = 2`. What changed? Compute the sample standard deviation.
- (3) Create object `x3` by setting `sd = 10`. What changed? Compute the sample standard deviation.
- (4) Create object `x4` by setting `sd = 0`. What happened now? Compute the sample standard deviation.

`@hint`
- Create the objects `x2`, `x3` and `x4` similarily to `x1` by calling `simulate()` and modifying the `sd` argument.
- Compute the standard deviations of the new object on different lines, similarily to the example.

`@pre_exercise_code`
```{r}

simulate <- function(sd = NULL, n = 100, mean = 0) {
  
  if(is.null(sd)) stop("Please provide the argument 'sd'")
  
  set.seed(666)
  
  x <- data.frame(x = rnorm(n = n, mean = mean, sd = sd), y = 1)
  sample_mean <- round(mean(x$x),1)
  
  plot(x, 
       xlim = c(-10, 10), 
       ylab="", xlab = "", yaxt = "n",
       main = "Values of x",
       sub = paste("sample size =",n,"\nsample mean = ",sample_mean),
      type = "n")
  abline(v = sample_mean, col = "red")
  points(x, pch = "|", col = "blue")
  
  return(x$x)
}

```

`@sample_code`
```{r}
# simulate() is available to draw samples from an artificial population

# Simulate data and visualize the distances to the mean (red line).
x1 <- simulate(sd = 1)  # (1)
x2 <-  # (2)
  # (3)
  # (4)

# Compute the average distances from the mean in the sample.
sd(x1)  # (1)
  # (2)
  # (3)
  # (4)

```

`@solution`
```{r}
# simulate() is available to draw samples from an artificial population

# Simulate data and visualize the distances to the mean (red line).
x1 <- simulate(sd = 1)    # (1)
x2 <- simulate(sd = 2)    # (2)
x3 <- simulate(sd = 10)   # (3)
x4 <- simulate(sd = 0)    # (4)

# Compute the average distances from the mean in the sample
sd(x1)  # (1)
sd(x2)  # (2)
sd(x3)  # (3)
sd(x4)  # (4)

```

`@sct`
```{r}

test_object("x1")
test_object("x2")
test_object("x3")
test_object("x4")

test_output_contains("sd(x1)")
test_output_contains("sd(x2)")
test_output_contains("sd(x3)")
test_output_contains("sd(x4)")

test_error()
success_msg("Awsome work!")

```

---

## Clustering

```yaml
type: NormalExercise
key: 989f75fe50
lang: r
xp: 100
skills: 1
```

[Clustering](https://en.wikipedia.org/wiki/Cluster_analysis) is the task of grouping a set of objects in such a way that objects in the same group (called a cluster) are more similar to each other than to those in other groups. [K-means](https://en.wikibooks.org/wiki/Data_Mining_Algorithms_In_R/Clustering/K-Means#Algorithm) is a popular algorithm that performs clustering. With k-means, data points that share the same closest *centroid* are concidered similar. k centroids form k clusters around them.  

The centroids are found in such a way that the combined distance from each data point to it's centroid is the smallest possible. In a simple case this is identical to saying that the standard deviations inside the clusters is the smallest possible.

`@instructions`
- Modify the call to `kmeans()` to group the student heights to two group by using two centroids (centers).
- Print out `km$centers`. Can you find any interpretations?
- Create a new colum `height_cluster` in `students2014` that stores the new clusters of the students. Look at the first 6 rows of `students2014` with `head()`.
- Draw the plot of the student heights, where the students are separated and colored by sex. Adjust the `col` argument of `plot()` and instead color the points by the height_clusters.

`@hint`
- Define the number of clusters using the `centers` argument of `kmeans()`.
- Replace `$sukup` with `$height_cluster` for the `col` argument of `plot()`.

`@pre_exercise_code`
```{r}
# load data from web
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
# recode kone variables missing values as factor levels
students2014$kone <- addNA(students2014$kone)
# keep only rows without missing values
students2014 <- students2014[complete.cases(students2014),]
# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)

```

`@sample_code`
```{r}
# students2014 is available
height <- students2014$pituus

# Use k-means to group the students by height
km <- kmeans(height, centers = 1)

# Print out the centroid values (centers)
km$centers

# Assign the clusters to the student data
students2014$height_cluster <- km$cluster

# Look at the first six values of students2014 with head()


# Trick to plot the student heights separated by sex.
sex <- jitter(as.numeric(students2014$sukup))
plot(sex, height, col = students2014$sukup)

```

`@solution`
```{r}
# students2014 is available
height <- students2014$pituus

# Use k-means to group the students by height
km <- kmeans(height, centers = 2)

# Print out the centroid values (centers)
km$centers

# Assign the clusters to the student data
students2014$height_cluster <- km$cluster

# Look at the first six values of students2014 with head()
head(students2014)

# Trick to plot the student heights separated by sex.
sex <- jitter(as.numeric(students2014$sukup))
plot(sex, height, col = students2014$height_cluster)

```

`@sct`
```{r}

test_function("kmeans",args=c("x","centers"))
test_object("students2014")
test_function("head", args = "x", not_called_msg = "Please use `head()` on the data.frame `students2014`.")
test_function("plot", args = "col")

test_error()
success_msg("Very nice work!")

```

---

## Guess the correlation

```yaml
type: NormalExercise
key: 1d52143e10
lang: r
xp: 100
skills: 1
```

Correlation measures the strength of the *linear* relationship between two variables. The easiest way to check if the relationship between two variables might be  interpret as linear is to visualize the relationship with a scatter plot. 

It is then possible to make a reasonable guess of the strength of the linear relationship. In order for you to practise this skill, we have build a game you can play.

`@instructions`
- Use `simulate()` to draw a scatter plot using simulated data. The variables can have linear and nonlinear relationships. 
- Try to guess what the correlation might be and save your guess to `guess`.
- Use `show_correlation()` to check how close you got.
- You can experiment as many times as you wish and the simulation will be different each time. Feel free to change `n`, the number of simulated values.

`@hint`
- When you are done experimenting, simply click 'Submit Answer' to move forward.

`@pre_exercise_code`
```{r}
# simulate and plot correlated data
simulate <- function(n = 100) {
  
  if(n > 1e5) stop("That n is quite large. Please choose a smaller n")
  
  # degree (d) of the polynomial
  degree <- sample(1:3, 1)
  # x values
  x <- rnorm(n)
  
  # each x^d multiplied by it's coefficient
  # the coefficients are sampled from U(-1, 1)
  X <- sapply(1:degree, function(d) {
    b <- runif(1, min = -1, max = 1)
    b * x^d
  })
  
  # linear equation y = b_1 * x^1 + .. + b_d * x^d + e
  y <- rowSums(X) + rnorm(n)
  # scatterplot
  plot(x, y, col = "grey40", type = "p", pch= 19, xlab="", ylab="", main = "Guess the correlation")
  
  return(list(x = x, y = y))
}

show_correlation <- function(df, guess) {
  x <- df$x
  y <- df$y
  correlation <- round(cor(x, y), 2)
  diff <- round(abs(correlation - guess),2)
  if(diff < 0.1) fb <- "you're a wizard!"
  else if(diff < 0.2) fb <- "awsome!"
  else if(diff < 0.3) fb <- "nice try!"
  else if(diff < 0.4) fb <- "not too bad."
  else if(diff >= 0.4) fb <- "try again?"
  plot(x, y, type = "p", col = "grey40", pch= 19, xlab="", ylab="")
  title(main = paste0("Results", "\n", 
                      "guess: ",guess," ",
                      "actual: ",correlation, "\n",
                      diff," difference, ",fb))
  abline(lm(y~x), col ="red")
}

```

`@sample_code`
```{r}
# Functions simulate() and show_correlation() are available

# Simulate n values and draw a scatterplot
data <- simulate(n = 100)

# What is the correlation?
guess <- 0.00

# Show the linear relationship and correlation
show_correlation(data, guess)

```

`@solution`
```{r}
# Functions simulate() and show_correlation() are available

# Simulate n values and draw a scatterplot
data <- simulate(n = 100)

# What is the correlation?
guess <- 0.00

# Show the linear relationship and correlation
show_correlation(data, guess)

```

`@sct`
```{r}

test_error()
success_msg("Thanks for playing!")

```
