---
title       : Hypothesis testing
description : UNDER CONSTRUCTION

--- type:MultipleChoiceExercise lang:r xp:100 skills:1 key:ee77c10ab2

## Statistical hypothesis testing

In [statistical hypothesis testing](https://en.wikipedia.org/wiki/Statistical_hypothesis_testing) we want to test if our claim or something we assume is supported by the data. So we set our [*null hypothesis*](https://en.wikipedia.org/wiki/Null_hypothesis) and our [*alternative hypothesis*](https://en.wikipedia.org/wiki/Alternative_hypothesis), and see what the numbers tell us about them.  



*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}

```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:c033eeaa78

## Meet the tests! (1)

After setting the hypotheses, we can look at the data and calculate the test statistic. What test statistic to use, depends on the test you are doing. With the test statistics, you can calculate the (famous) p-value. The p-value tells you something about the hypotheses and it can be used as a guide line. The p-value and other knowledge of the field you are testing, you can reject or not reject the null hypothesis. 
One of the tests is to test whether some variable is distributed as we assume. Often statistical assumptions hold if the variables are normally distributed. 

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
learning2014 <-  read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/learning2014.txt", sep = "\t", header = TRUE)
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:9b63d0cb6a

## Meet the tests! (2)

T-test is very common statistical test. It can be used to test if the means of two groups differ from each other statistically. So it has hypotheses 

$$H_0: mean_1 = mean_2
H_1: mean_1 \neq mean_2$$

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# Separate dataset by gender
women <- subset(learning2014, gender == 'N')
men <- subset(learning2014, gender == 'M')

# T-test
t.test(women$deep, men$deep)
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:0b42a1089b

## Peculiar p-values

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:8dfc56d628

## (Null hypothesis:) which way to go?!

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

