---
title       : Data types and variable types
description : Quick overview to R data and variable types. What are the objects? You are the subject.

--- type:NormalExercise lang:r xp:100 skills:1 key:8718babe2e
## Data frames

In R, the basic idea is to use *functions* to operate on objects. The most common object to operate on is the `data frame` object, which stores data in tabular form.  

The columns in data frames are vectors. Data frames can contain multiple vectors with different data types. In this chapter you will learn more about vectors and data types.

Here you will get to apply some functions that are helpful for inspecting data: `head()` and `str()`. You will also learn how to access the columns of a data frame using the $-sign.  

*** =instructions
- Execute the function `head()` and see the results
- Inspect the data types in the `students2014` data with the function `str()`
- Use the $-sign to access the students shoe size (`kenka`) in `students2014` data

*** =hint
- The data is stored in object `students2014`

*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)

# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]

# recode kone -variable NA values as factor levels
students2014$kone <- addNA(students2014$kone)

# choose rows without missing values
students2014 <- students2014[complete.cases(students2014),]

# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```

*** =sample_code
```{r}
# students2014 is available

# Let's view the first 6 students from the course participants data
head(students2014)

# View the structure of students2014
str(students2014)

# You can access the variables in the students2014 data by using the $-sign
students2014$ika

# Access the shoe sizes of the students


```

*** =solution
```{r}
# students2014 is available

# Let's view the first 6 students from the course participants data
head(students2014)

# View the structure of students2014
str(students2014)

# You can access the variables in the students2014 data by using the $-sign
students2014$ika

# Access the shoe sizes of the students
students2014$kenka
```

*** =sct
```{r}
test_output_contains("students2014$kenka", incorrect_msg = "Did you use the $-sign to access the shoe size variable?")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:e4360dcfec
## Data types (1)

In this section you learn about the data types in R. You will also get familiar with vectors and their usage. What does data types and vectors have to do with each other? A lot, you'll see! 

There are different data types in R. The most commonly used are:

- Numbers and decimal values are called **numeric**
- Text or character values are called **characters**
- Boolean values (`TRUE` or `FALSE`) are called **logical**

When creating character values, one needs to use quotation marks (both `' '` and `" "` will work). For example `"Statistics<3"` would be a character value in R. Numeric and boolean values do not need quotation marks. 

You can use the function `str()` (str is short for structure) to find out what data type the object is.

*** =instructions
- Create character object `my_initials` by assigning your initials in it.
- Assign the number 10.5 in `my_numeric_object`. Note that the decimal separator in R is dot instead of comma. 
- Create the object `my_boolean` by executing the line. Note that when creating boolean values, no quotation marks are needed.
- Change the value of `my_boolean` to `TRUE`
- Inspect the data types with function `str()`.

*** =hint
- When creating a character object, remember to use quotation marks.
- Dot is the decimal mark in R:  `2.5` is decimal number but `2,5` means a different thing!
- Remember that `Ctrl + Enter` executes a single row.
- You can use `str()` by putting the object name inside the parenthesis.

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# Create objects with different data types
my_initials <-
my_numeric_object <-
my_boolean <- FALSE

# See how the objects look with the function str()


```

*** =solution
```{r}
# Create objects with different data types
my_initials <- "A.B.C."
my_numeric_object <- 10.5
my_boolean <- TRUE

# See how the objects look with the function str()
str(my_initials)
str(my_numeric_object)
str(my_boolean)
```

*** =sct
```{r}
test_object("my_initials", eval=FALSE, incorrect_msg = "Did you assign your initials to `my_initials`?")
test_object("my_boolean", incorrect_msg="Did you change the value of `my_boolean`?")
test_object("my_numeric_object", incorrect_msg="Did you assign correct value to  `my_numeric`?")

test_output_contains("str(my_boolean)", incorrect_msg = "Did you use `str()` on `my_boolean`?")
test_output_contains("str(my_initials)", incorrect_msg = "Did you use `str()` on `my_initials`?")
test_output_contains("str(my_numeric_object)", incorrect_msg = "Did you use `str()` on `my_numeric_object`?")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("You're making great progress, well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:9ee1ba578e
## Data types (2)

One of R's data types is called factor. Factors contain numbers, but for every number there is a character value associated to it. A variable measured by the [Likert scale](https://en.wikipedia.org/wiki/Likert_scale) could for example be coded as a factor: the values are numbers from 1 to 5, but the numbers also have words attached to them such as:

- "Strongly disagree" = 1
- "Strongly agree" = 5

In R is is also possible to create factors from numerical objects using the function `cut()`, which splits a variable into categories. It has a mandatory argument `breaks`, which defines the cutting points of the categories.  

*** =instructions
- Create objects `student_age` and `student_age_cut`
- use `str()` to inspect `student_age_cut`
- Use `head()` to print the first six values of `student_age`
- Use `head()` to print the first six values of `student_age_cut`


*** =hint
- Give the object `student_age_cut` as the first argument to `str()` and `head()`.

*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)

# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]

# recode kone -variable NA values as factor levels
students2014$kone <- addNA(students2014$kone)

# choose rows without missing values
students2014 <- students2014[complete.cases(students2014),]

# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```

*** =sample_code
```{r}
# students2014 is available

# Let's access the (numerical) age vector of the students2014 and save it to another object.
student_age <- students2014$ika

# Let's use the cut()-function to cut the variable in categories
student_age_cut <- cut(student_age, breaks = c(15, 25, 35, 100))

# Inspect the data type of student_age_cut


# Print the first 6 elements of student_age and student_age_cut



```

*** =solution
```{r}
# students2014 is available

# Let's access the (numerical) age vector of the students2014 and save it to another object.
student_age <- students2014$ika

# Let's use the cut()-function to cut the variable in categories
student_age_cut <- cut(student_age, breaks = c(15, 25, 35, 100))

# Inspect the data type of student_age_cut
str(student_age_cut)

# Print the first 6 elements of student_age and student_age_cut
head(student_age)
head(student_age_cut)


```

*** =sct
```{r}

test_function("str", args="object", incorrect_msg = "Did you use the `str()` to look at the object 'student_age_cut'?")

test_output_contains("head(student_age", incorrect_msg = "Did you use `head()` and look at the first six elements of `student_age`?")
test_output_contains("head(student_age_cut)", incorrect_msg = "Did you use `head()` to look at the first six elements of `student_age_cut`?")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:7cc56d2158
## Vectors

A vector is an object containing multiple elements of the same data type. Vectors are quite important in R. The columns of a data frame (such as `student2014`) are vectors.

You can also create vectors by combining values of the same data type, using the function `c()`. 

In this exercise you will learn how to create your own vector.

*** =instructions
- See and execute the examples on how to create a vector with `c()`.
- Create a character vector `my_vector`.
- Create a vector `boolean_vector` with boolean values `TRUE, FALSE, TRUE, FALSE`.
- Print the contents of `boolean_vector`.

*** =hint
- When creating boolean values, no quotation marks are needed.
- You can print the contents of an object by typing it's name.


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# Here we create a vector with numeric elements
c(2, 3, 4, 5)

# Another one
c(0.1, 0.2, 5.84, 0.7)

# Let's create one with characters and save it to a object
my_vector <- c("a", "b", "c", "d")

# Create boolean vector and print it


```

*** =solution
```{r}
# Here we create a vector with numeric elements
c(2, 3, 4, 5)

# Another one
c(0.1, 0.2, 5.84, 0.7)

# Let's create one with characters and save it to a object
my_vector <- c("a", "b", "c", "d")

# Create boolean vector and print it
boolean_vector <- c(TRUE, FALSE, TRUE, FALSE)
boolean_vector

```

*** =sct
```{r}

test_object("boolean_vector", incorrect_msg = "Did you create `boolean_vector` with same values as instructed?")
test_output_contains("boolean_vector", incorrect_msg = "Did you print the contents of `boolean_vector`?")
test_error()

# Final message the student will see upon completing the exercise
success_msg("Great work!")

```

--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:3c0486635f
## Data types and measurement scales

Data types and measurement scales are somewhat related. When asking simple "Yes / No " -questions boolean values are useful but you would not use them for measuring heights. Here is a table for the data types and measuring scales:

| Scale          | Numeric | Character | Boolean | Factor |
|----------------|---------|-----------|---------|--------|
| Categorical  |    x    |     x     |    x    |    x   |
| Ordinal        |    x    |     x     |         |    x   |
| Measurement    |    x    |           |         |        |


Use the R Console to inspect the `students2014` data with the function `str()` and select the correct answer from below. The data is already loaded for you. Remember that in the console `Enter` executes lines.

Choose one.  

*** =instructions
- Variable `kone` is from ordinal scale.  
- Variable `ika` and `pituus` are character values.  
- You can use boolean data type for variables `ika` and `pituus`.  
- Variable `kone` is from categorical scale.  
- It is meaningful to use factors for measurement scale.  


*** =hint
- There is no natural order in categorical scale. 
- If you can put the values in a meaningful order, then it's ordinal scale. 
- Using factors probably loses its purpose when you have a lot of different values.
- Character values have quotations marks. `chr` is short for character. 

*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)

# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]

# recode kone -variable NA values as factor levels
students2014$kone <- addNA(students2014$kone)

# choose rows without missing values
students2014 <- students2014[complete.cases(students2014),]

# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```

*** =sct
```{r}
msg1 = "Sorry, I know you feel like Mac is better than Windows put in this case there is no order. Try again!"
msg2 = "I would use numbers to measure height and age. Try again!"
msg3 = "Maybe you can do that but it makes no sense. Guess again!"
msg4 = "Well done. Proceed to the next exercise"
msg5 = "Well technically you CAN do that, but you'll have as many factors as there are different numbers. Why not just use numerical values? Guess again!"

test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4,msg5))

# Final message the student will see upon completing the exercise
success_msg("Nicely done!")

```


