---
title       : Exploring variation and dependence
description : Variation and dependence are at the heart of Statistics. In fact, without variation the discipline would cease to exist. Correlation is not causation, but why?

--- type:NormalExercise lang:r xp:100 skills:1, 3 key:2978bc79da
## summary() statistics

Some functions in R are very generic. They can take different objects as their first argument and they seem to magically do exactly what you'd hope they do in all situations.

One such example is the `summary()` function.

*** =instructions
- Execute the code that creates a summary of the `variable` in `df`
- Then create a summary of `df`
- `summary()` behaves differently because `variable` and `df` are different kinds of objects.

*** =hint
- Get the summary of an object by typing summary(object)

*** =pre_exercise_code
```{r}
# pec
```


*** =sample_code
```{r}

# Create a summary of var
summary(df$var)

# Create a summary of df



```

*** =solution
```{r}

# Create a summary of var in df
summary(df$var)

# Create a summary of df
summary(df)


```

*** =sct
```{r}
# tests. 
test_output_contains("summary(df)", incorrect_msg="Please type in the code to create a summary of `df`")

test_error()
success_msg("Great job!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:74f8535355

--- type:NormalExercise lang:r xp:150 skills:1 key:c3d62c54f5

## Box plots

One very good way to visualize the most important summary statistics of a variable is by drawing box plots. A box plot visualizes the 25th, 50th and 75th percentiles (the box), the typical range (the whiskers) and the outliers (the dots) of a variable 

The whiskers extending from the box can be computed by several techniques. The default in R is to extend them to reach to a data point that is no more than 1.5*IQR away from the box, where IQR is the inter quartile range defined as  

`IQR = 75th percentile - 25th percentile`  

Values outside the whiskers can be considered as outliers, unusually distant observations. For more information on IQR, see <a target="_blank" href ="https://en.wikipedia.org/wiki/Interquartile_range"> wikipedia</a>

*** =instructions
- Draw a `boxplot()` of `var`.
- Use help() or `?` to find out which argument controls the label of the y axis and the main title.
- Add a suitable label to the y axis and a main title of your choice.

*** =hint
- hint1

*** =pre_exercise_code
```{r}
# pec
```


*** =sample_code
```{r}
# data frame is available

# See the help page of boxplot()
?boxplot

# Draw a boxplot of var. Add a y axis label and a main title.



```

*** =solution
```{r}
# Draw a boxplot of var. Add a y axis label and a main title.
boxplot(var, ylab = "", main = "")


```

*** =sct
```{r}


test_function("boxplot", args="x")

# test error
test_error()


# Final message the student will see upon completing the exercise
success_msg("Awsome work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:8a451ced6e

## Exercise 3

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:a8a509dfd2

## Exercise 4

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:d66995b6b4

## Exercise 5

Under construction.

*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# sample code here
```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

