---
title       : Data types and variable types
description : Quick overview to R data and variable types. What are the objects? You are the subject.

--- type:NormalExercise lang:r xp:100 skills:1 key:e4360dcfec
## Data types in R (1)

There are different data types in R. The most commonly used are:

- Numbers and decimal values are called **numeric**
- Text or character values are called **characters**
- Boolean values (`TRUE` or `FALSE`) are called **logical**

You can use the function `str()` (str is short for structure) to find out what data type the object is.

*** =instructions
- Create character object `my_initials` by assigning your initials in it.
- Assign the number 10.5 in `my_numeric_object`. Note that the decimal separator in R is dot instead of comma. 
- Create the object `my_boolean` by executing the line. Note that when creating boolean values, no quotation marks are needed.
- Change the value of `my_boolean` to `TRUE`
- Inspect the data types with function `str()`.

*** =hint
- When creating a character object, remember to use quotation marks.
- Dot is the decimal mark in R:  `2.5` is decimal number but `2,5` means a different thing!
- Remember that `Ctrl + Enter` executes a single row.
- You can use `str()` by putting the object name inside the parenthesis.

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# Create objects with different data types
my_initials <-
my_numeric_object <-
my_boolean <- FALSE

# Change the value of my_boolean to TRUE


# See how the objects look with the function str()




```

*** =solution
```{r}
# Create objects with different data types
my_initials <- "A.B.C."
my_numeric_object <- 10.5
my_boolean <- FALSE

# Change the value of my_boolean
my_boolean <- TRUE

# See how the objects look with the function str()
str(my_initials)
str(my_numeric_object)
str(my_boolean)
```

*** =sct
```{r}
test_student_typed("my_boolean <- TRUE", not_typed_msg="Did you change the value of `my_boolean`?")
test_object("my_initials", eval=FALSE, incorrect_msg = "Did you assign your initials to `my_initials`?")
test_student_typed("my_numeric_object <- 10.5", not_typed_msg="Did you assign correct value to  `my_numeric`?")

test_function("str", incorrect_msg = "Did you use the str() to look at the data types?")

test_output_contains("str(my_boolean)", incorrect_msg = "Did you use `str()` to `my_boolean`?")
test_output_contains("str(my_initials)", incorrect_msg = "Did you use `str()` to `my_initials`?")
test_output_contains("str(my_numeric_object)", incorrect_msg = "Did you use `str()` to `my_numeric_object`?")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("You're making progress, well done!")
```

--- type:NormalExercise lang:r xp:100 skills:1 key:9ee1ba578e
## Data types in R (2)

One of R's data types is called factor. It turns (for example) character objects to objects with a numerical encoding and creates corresponding label names.

The function `cut()` can be used to create categories from numerical objects. The function produces a factor object and by default creates label names or levels. The function `cut()` also has a mandatory argument `breaks`, where the break points of the object are assigned.

NA/NULL?

*** =instructions
- Create objects `student_age` and `student_age_cut`
- Print the contents of the `student_age_cut`
- use `str()` to inspect the object `student_age_cut`

*** =hint
- Functions do not work correctly without mandatory arguments.
- In the `cut()` function the argument `breaks` 

*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
```

*** =sample_code
```{r}
# students2014 is available

# Let's access the (numerical) age vector of the students2014 and save it to another object.
student_age <- students2014$ika

# Let's use the cut()-function to cut the variable in categories
student_age_cut <- cut(student_age, breaks = c(15, 25, 35, 45, 55, 65, 75, 85))

# print the contains and inspect the data type of student_age_cut



```

*** =solution
```{r}
# students2014 is available

# Let's access the (numerical) age vector of the students2014 and save it to another object.
student_age <- students2014$ika

# Let's use the cut()-function to cut the variable in categories
student_age_cut <- cut(student_age, breaks = c(15, 25, 35, 45, 55, 65, 75, 85))

# print and inspect the data type of student_age_cut
student_age_cut
str(student_age_cut)
```

*** =sct
```{r}
test_student_typed("student_age <- students2014$ika", not_typed_msg="Did you create the `student_age` object?")
test_student_typed("student_age_cut <- cut(student_age, breaks = c(15, 25, 35, 45, 55, 65, 75, 85))", not_typed_msg="Did you create the `student_age_cut` object?")

test_function("str", incorrect_msg = "Did you use the str() to look at the factor data type?")
test_output_contains("str(student_age_cut)", incorrect_msg = "Did you use `str()` to `student_age_cut`?")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:7cc56d2158
## Vectors (1)

You can combine values of the same data type to a **vector**, using the function `c()`. 

A vector is an object containing multiple elements. Vectors are quite important in R. The columns of a data frame (such as `student2014`) are vectors.

In this execrice you will learn how to create your own vector.

*** =instructions
- See and execute the examples on how to create a vector with `c()`.
- Save character vector to `my_vector`.
- Create vector `boolean_vector` with boolean values `TRUE, FALSE, TRUE, FALSE`.

*** =hint
- When creating boolean values, no quotation marks are needed.


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}

# Here we create a vector with numeric elements
c(2, 3, 4, 5)

# Another one
c(0.1, 0.2, 5.84, 0.7)

# Let's create one with characters and save it to a object
my_vector <- c('a', 'b', 'c', 'd')

# Create boolean vector


```

*** =solution
```{r}

# Here we create a vector with numeric elements
c(2, 3, 4, 5)

# Another one
c(0.1, 0.2, 5.84, 0.7)

# Let's create one with characters and save it to a object
my_vector <- c('a', 'b', 'c', 'd')

# Create boolean vector
boolean_vector <- c(TRUE, FALSE, TRUE, FALSE)

```

*** =sct
```{r}
test_output_contains("c(2, 3, 4, 5)", incorrect_msg = "Did you execute all the lines?")
test_output_contains("c(0.1, 0.2, 5.84, 0.7)", incorrect_msg = "Did you execute all the lines?")

test_object("my_vector <- c('a', 'b', 'c', 'd')", incorrect_msg = "Did you execute `my_vector`?")
test_object("boolean_vector <- c(TRUE, FALSE, TRUE, FALSE)", incorrect_msg = "Did you create `boolean_vector` with same values as instructed?")
test_error()

# Final message the student will see upon completing the exercise
success_msg("Great work!")

```
--- type:NormalExercise lang:r xp:100 skills:1 key:12a4e8cd24
## Vectors (2)

*** =instructions
- 
*** =hint
-

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}


```

*** =solution
```{r}

```

*** =sct
```{r}
test_output_contains("c(2, 3, 4, 5)", incorrect_msg = "Did you execute all the lines?")
test_output_contains("c(0.1, 0.2, 5.84, 0.7)", incorrect_msg = "Did you execute all the lines?")

test_object("my_vector <- c('a', 'b', 'c', 'd')", incorrect_msg = "Did you execute `my_vector`?")
test_object("boolean_vector <- c(TRUE, FALSE, TRUE, FALSE)", incorrect_msg = "Did you create `boolean_vector` with same values as instructed?")
test_error()

# Final message the student will see upon completing the exercise
success_msg("Great work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:17c91c6d3c
## Data types and functions

Not all data types do well with certain functions. 

One cannot for example use character vectors when the function needs a numeric one. 

When using the wrong data type, R produces an error stating what is wrong. 

*** =instructions
- Create `vector`.
- Try to calculate the mean of `vector` and see what happens. 
- Correct the vector and calculate the mean so that you do not get any errors.

*** =hint
- You can use `str()` to check the data types.
- `chr` is short for character.
- When creating a character vector, one needs to use quotation marks. 


*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# Let's create a vector
vector <- c(2, 3, '1', 5)

# Calculate the mean of vector
mean(vector)
  
```

*** =solution
```{r}

# Let's create a vector
vector <- c(2, 3, '1', 5)

# Calculate the mean of vector
mean(vector)

```

*** =sct
```{r}
test_object("vector <- c(2, 3, 1, 5)", incorrect_msg = "Did you correct the `vector`?")
test_function("mean", args = "x")
test_error()

# Final message the student will see upon completing the exercise
success_msg("Creating vectors like a pro. Great work!")

```
--- type:NormalExercise lang:r xp:100 skills:1 key:8718babe2e
## Data frames

The most common object to operate on is the `data frame` object, which stores data in tabular form. The columns in data frames are vectors. Data frames can contain multiple vectors with different data types.

In R, the basic idea is to use *functions* to operate on objects. Here you will get to apply some functions that are helpful for inspecting data: `head()` and `str()`.

You will also learn how to access the columns of a data frame using the $-sign.

*** =instructions
- Execute the function `head()` and see the results
- Inspect the data types in the `students2014` data with the function `str()`
- Use the $-sign to access the students shoe size (`kenka`) in `students2014` data

*** =hint
- The data is stored in object `students2014`

*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
```

*** =sample_code
```{r}
# students2014 is available

# Let's view the first 6 students from the course participants data
head(students2014)

# View the data types in the students2014 data


# You can access the variables in the students2014 data by using the $-sign. For example, you get the age variable of the students with
students2014$ika

# Access the shoe sizes of the students


```

*** =solution
```{r}
# Let's view the first 6 students from the course participants data
head(students2014)

# View the data types in the JYT data
str(students2014)

# You can access the variables in the students2014 data by using the $-sign. For example, you get the age variable of the students with
students2014$ika

# Access the shoe sizes of the students
students2014$kenka
```

*** =sct
```{r}
test_function("head", incorrect_msg = "Did you use the head() function?")
test_function("str", incorrect_msg = "Did you use the str() function?")
test_output_contains("students2014$kenka", incorrect_msg = "Did you use the $-sign to access the shoe size variable?")

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```

--- type:NormalExercise lang:r xp:100 skills:1 key:3c0486635f
## Data types and variable types/scales (?)


| Scale          | Numeric | Character | Boolean | Factor |
|----------------|:---------:|:-----------:|:---------:|:--------:|
| Classification |    x    |     x     |    x    |    x   |
| Ordinal        |    x    |     x     |         |    x   |
| Measurement    |    x    |           |         |        |

*** =instructions
- 

*** =hint
- 

*** =pre_exercise_code
```{r}
col1 <- rgb(229, 5, 58, max = 255)
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)

barplot(table(students2014$kone), sub="(M=Mac, W=Windows, L=Linux)", col=col1)
barplot((table(students2014$kenka)), main="Shoe sizes", col=col1)
hist(students2014$ika, breaks="scott", main="Age", col=col1)
student_age_cut <- cut(students2014$ika, breaks = c(15, 25, 35, 45, 55, 65, 75, 85))
plot(student_age_cut, main="student_age_cut", col=col1)
```

*** =sample_code
```{r}
# Classification
barplot(table(students2014$kone))

# Ordinal scale
barplot((table(students2014$kenka)))

# Measurement
hist(students2014$ika, breaks = "scott")

# From measurement to ordinal scale
student_age_cut <- cut(students2014$ika, breaks = c(15, 25, 35, 45, 55, 65, 75, 85))
plot(student_age_cut, main = "student_age_cut")
```

*** =solution
```{r}
```

*** =sct
```{r}
# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


