---
title       : Working across the data
description : Cross-tabulations let you explore dependencies hidden deep within discrete variables.

--- type:NormalExercise lang:r xp:100 skills:1 key:e1b927030f

## Combining variables

Often in a survey there are many questions that measure the same phenomenon. The phenomenon could be for example trustfulness or prejudice of some sort. Single survey questions are nice, but they do not help us to see the big picture. To get a better outlook of a phenomenon, we can combain multiple variables into one. This can be done by summing up the one person's answers. We can also combain the variables by taking the means of one person's answers. 

Next we learn to do that by ourselves with a little toy example. See the sample code!


*** =instructions
- In our toy example, the dataset is called `data`. Print out its contains.
- On the rows are the dudes that answered in the survey. The columns are the items (questions). Let's say that all the items measure the same phenomenon.


*** =hint
- See how the compaining of variables is done from the sample code.


*** =pre_exercise_code
```{r}
Emma <- c(2, 3, 2)
Pihla <- c(4, 5, 5)
Marianne <- c(1, 2, 1)
Liisi <- c(2, 3, 3)
Petteri <- c(4, 1, 3)
data<-data.frame(rbind(Emma, Pihla, Marianne, Liisi, Petteri))
colnames(data)<-c("Item_1", "Item_2", "Item_3")
```


*** =sample_code
```{r}
# In our toy example, the dataset is called 'data'. Print out its contains.
data

# Calculate the sums for every person for 'summamuuttuja_1'. Then create summamuuttuja_2
data$summamuuttuja_1  <- rowSums(data)
data$summamuuttuja_2  <- data$summamuuttuja_1 / 3

# Print the data again.
```


*** =solution
```{r}
# In our toy example, the dataset is called 'data'. Print out its contains.
data

# Calculate the means for every person. 
data$summamuuttuja <- rowMeans(data)

```


*** =sct
```{r}

test_error()

success_msg("Summing those variables up pretty nicely!")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:24b2372fff

## Tables

Cross tabulation is important form of presenting information. From tables one can see details that plots and pictures do not necessarily convey. Let's do couple of simple tables from the `students2014` data.

Maybe the most basic functions to do cross tabulations in R are `table()` and `prop.table()`. With the `table()` function you can create freguency tables and with the function `prop.table()` proportional tables are created. The `prop.table()` function needs a table object as an argument. Look the examples to see how tables are created.

*** =instructions
- Create the object `gender_tbl` and print its contents to see how the table looks like.
- Create proportional table by using the `gender_tbl` object and see the results.
- Use the `table()` and `prop.table()` functions to create another proportional table of the variable `kone` from the `students2014` dataset.

*** =hint
- Remember to use the $-sign to access variables of the `students2014` dataset.
- Function `prop.table()` needs a product of the function `table()` as an argument to work correctly. 


*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)

# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]

# recode kone -variable NA values as factor levels
students2014$kone <- addNA(students2014$kone)

# choose rows without missing values
students2014 <- students2014[complete.cases(students2014),]

# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```


*** =sample_code
```{r}
# students2014 is available

# Create a table of 'sukup' in 'students2014'
gender_tbl <- table(students2014$sukup)

# Print the table
gender_tbl

# Create a proportional table of 'sukup' in 'students2014'
prop.table(gender_tbl)

# Create a proportional table of 'kone' in 'students2014'



```


*** =solution
```{r}
# students2014 is available

# Create a table of 'sukup' in 'students2014'
gender_tbl <- table(students2014$sukup)

# Print the table
gender_tbl

# Create a proportional table of 'sukup' in 'students2014'
prop.table(gender_tbl)

# Create a proportional table of 'kone' in 'students2014'
prop.table(table(students2014$kone))

```


*** =sct
```{r}
# submission correctness tests

test_function("prop.table", args=c("x"), incorrect_msg = "Did you use prop.table on a table of `kone`?")

test_error()

success_msg("Making progress, good work!")

```


--- type:MultipleChoiceExercise lang:r xp:50 skills:1 key:6fad9de433

## Metadata of students2014

The `students2014` data is used in many of the exercises. The dataset was collected in 2014 from the students of the Faculty of Social Sciences in the University of Helsinki. This [questionnaire](http://www.helsinki.fi/~kvehkala/JYTmooc/ASSIST-2014-FI-0-20140903.pdf) was used to collect the information.

To use the collected data, you will have to put it in a digital form. The `students2014` data was saved manually. When saving the data it is often useful to use short descriptive names for the questions (variable names in the dataset).

The meta file for the `students2014` can be seen [here](http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-meta.txt). Note that in the exercises we have used only few variables. From the metafile you can check what the questionnaire's answer options were and some other information.

Open and read the meta file. What does `95` mean in the variable `opalku`?  

*** =instructions
- 95 is the number of persons who answered in the questionnaire  
- It means that the answerer has started his/hers studies in 1995  
- There were 95 questions in the questionnaire  
- It is a bluff: 95 is not mentioned in the meta file  
- The average of credits is 95 among the answerers  


*** =hint
- Did you open the [meta file](http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-meta.txt)?  
- If the link does not work, go to http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-meta.txt  


*** =pre_exercise_code
```{r}
open_meta <- function() browseURL("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-meta.txt")
```


*** =sct
```{r}
msg1 = "Sorry, that's not the right answer. Look again!"
msg2 = "You are totally right! Proceed to the next exercise"
msg3 = "Not quite right. Check again.!"
msg4 = "No it's there, look again!"
msg5 = "Sorry, that's not the right answer. Look again!"

test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4,msg5))

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:a7f8986e45

## Cross tabulations

Tables get more interesting when there is more than one variable. You can find fascinating information by doing simple cross tabulations on different data variables. 

The function `table()` can take more than one variable and produce more complex cross tabulations. When using `table()` with more than one variable, one needs to separate the variables with a comma (`,`). 

See how the `prop.table()` works with multiple variables from the sample code. Also you will learn useful function `addmargins()` which calculates the margins of the table.

*** =instructions
- Create object `gender_work_tbl` and see how the table of variable `sukup` and `toita` looks like. 
- Look at the proportional table of `gender_work_tbl`.
- Look what happens when you use the `addmargins()` function on `gender_work_tbl`.
- Create `gender_os_tbl` with variables `sukup` and `kone` from the `students2014`
- Print `gender_os_tbl` and add margins to it.

*** =hint
- Did you separate the variable names with comma (`,`) inside the `table()` function?
- Both `prop.table()` and `addmargins()` need a product of `table()` function as an argument to work correctly. 


*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)

# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]

# recode kone -variable NA values as factor levels
students2014$kone <- addNA(students2014$kone)

# choose rows without missing values
students2014 <- students2014[complete.cases(students2014),]

# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```


*** =sample_code
```{r}
# students2014 is available

# Create a table of 'sukup' and 'toita'. Print the table.
gender_work_tbl <- table(students2014$sukup, students2014$toita)


# Look at the proportional table and freguency table with margins.
prop.table(gender_work_tbl)
addmargins(gender_work_tbl)

# Create a table of 'sukup' and 'kone' from the 'students2014'
gender_os_tbl <-

# Print 'os_gender_tbl'

    
# Add margins to 'os_gender_tbl'
  
  
```


*** =solution
```{r}
# students2014 is available

# Create a table of 'sukup' and 'toita'. Print the table.
gender_work_tbl <- table(students2014$sukup, students2014$toita)


# Look at the proportional table and freguency table with margins.
prop.table(gender_work_tbl)
addmargins(gender_work_tbl)

# Create a table of 'sukup' and 'kone' from the 'students2014'
gender_os_tbl <- table(students2014$sukup, students2014$kone)

# Print 'os_gender_tbl'
gender_os_tbl
    
# Add margins to 'os_gender_tbl'
addmargins(gender_os_tbl)
    
```


*** =sct
```{r}
test_object("gender_os_tbl")
test_function("addmargins", args=c("A"), incorrect_msg = "Did you use the `addmargins()` function?")
test_output_contains("gender_os_tbl", incorrect_msg = "Did you print `gender_os_tbl`?")

test_error()

success_msg("Great! Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:242f06b715

## Let's (bar)plot that table

Now we know how to create a table and next we want to visualize it. The basic functions to use when plotting tables are 
`barplot()` and `plot()`. 

The `barplot()` function has many arguments, and it is often useful to see the help page of the function with `help(barplot)` or `?barplot`. We will pay special attention to the arguments `beside` and `legend.text`. You will notice the difference in plots, see the example code!

*** =instructions
- Execute objects `gender_os_tbl`.
- Plot barplot of `gender_os_tbl` when `beside` is `FALSE`.
- Change `beside` to `TRUE`. Do you see the difference?
- Do a better barplot of `gender_os_tbl` by adding a legend. You can add a legend with code `legend.text = rownames(gender_os_tbl)`. Note that the  `rownames()` is also a function, and it takes the table as an argument.


*** =hint
- In functions multiple arguments are separated with comma. 
- No quotation marks are needed with `FALSE` and `TRUE`. 


*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)

# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]

# recode kone -variable NA values as factor levels
students2014$kone <- addNA(students2014$kone)

# choose rows without missing values
students2014 <- students2014[complete.cases(students2014),]

# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```


*** =sample_code
```{r}
# students2014 is available

# Let's create 'gender_os_tbl'
gender_os_tbl <- table(students2014$sukup, students2014$kone)

# Barplot of 'gender_os_tbl'
barplot(gender_os_tbl, beside = FALSE)

```


*** =solution
```{r}
# students2014 is available

# Let's create 'gender_os_tbl'
gender_os_tbl <- table(students2014$sukup, students2014$kone)

# Barplot of 'gender_os_tbl'
barplot(gender_os_tbl, beside = TRUE, legend.text = rownames(gender_os_tbl))

```


*** =sct
```{r}

test_function("barplot", args=c("height", "beside"), not_called_msg = "Did you use arguments `beside` and `legend` in the barplot?")

test_error()

# Final message the student will see upon completing the exercise
success_msg("You are doing very well!")

```


--- type:NormalExercise lang:r xp:100 skills:1 key:954e6d31e8

## And then little more plotting

The `barplot()` function creates nice plots but sometimes, especially with proportional tables, you want to have a plot where you can see the ratios or proportions right away. R has default plot types for different kind of object, and when you use the function `plot()` where a table object is an argument, R produces a [mosaic plot](https://en.wikipedia.org/wiki/Mosaic_plot).

Mosaic plots are good when you want see the proportions of a table with a quick look. Unlike in barplots, the bars in the mosaic plot have different widths. With different bar widths, you immediately notice the size differences between the cells of the table. 

*** =instructions
- Execute `gender_os_prop`.
- Use the `plot()` function to create a plot of the `gender_os_prop`. Add a descriptive title to the plot.

*** =hint
- The `plot()` function can be used for many kinds of data. When there is a table object as an argument, `plot()` produces a mosaicplot.
- You can add a title to the plot with the `main` argument.

*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)

# keep a couple background variables
students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]

# recode kone -variable NA values as factor levels
students2014$kone <- addNA(students2014$kone)

# choose rows without missing values
students2014 <- students2014[complete.cases(students2014),]

# integers to numeric
students2014$ika <- as.numeric(students2014$ika)
students2014$pituus <- as.numeric(students2014$pituus)
students2014$kenka <- as.numeric(students2014$kenka)
```


*** =sample_code
```{r}
# students2014 is available

# Proportional table of 'sukup' and 'kone'
gender_os_prop <- prop.table(table(students2014$kone, students2014$sukup))

# Plot 'gender_os_prop'. Add a title. 


```


*** =solution
```{r}
# students2014 is available

# Proportional table of 'sukup' and 'kone'
gender_os_prop <- prop.table(table(students2014$sukup, students2014$kone))

# Plot 'gender_os_prop'. Add a title.
plot(gender_os_prop, main = "OS by gender")


```


*** =sct
```{r}
test_function("plot", args=c("main"), incorrect_msg = "Use the argument `main` in the `plot()` function to add a title.")

success_msg("You have learned so much! Well done!")

```
