--- type:NormalExercise lang:r xp:100 skills:1
## Brackets have your data [ under, control ]

It is quite often necessary to access only a part of a dataset. This is called *subsetting* and there are many ways of doing it. The best way will depend on the situation and  preference.  

We have already seen that the columns of a data frame (df) can be accessed via `$`. There is another, more general method: `[ ]`. The brackets are a way of accessing a part of any data structure in R. 

The way to use brackets is by giving the locations of the values you wish to access.


*** =instructions
- Study the examples on how to select parts of a data frame.
- Replace `NULL` so that the object `age_height` contains the third and fourth columns of students2014.
- Modify the code to print out the first five rows of `age_height`.


*** =hint
- Selecting multiple columns is very similar to selecting multiple rows.
- See the first example on how to select multiple rows.


*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
  
  # keep a couple background variables
  students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
  
  # recode kone -variable NA values as factor levels
  students2014$kone <- addNA(students2014$kone)
  
  # choose rows without missing values
  students2014 <- students2014[complete.cases(students2014),]

```

*** =sample_code
```{r}
# students2014 is available

# Choose the rows from 3 to 6 in the data
students2014[3:6, ]

# Look at the value in row 6, column 4
students2014[6, 4]

# Choose the whole column 3 and save it as an object
student_age <- students2014[, 3]

# Choose the third and fourth columns and save them as an object
age_height <- students2014[, NULL]

# Print out the first five rows of age_height
age_height[NULL, ]


```

*** =solution
```{r}
# students2014 is available

# Choose the rows from 3 to 6 in the data
students2014[3:6, ]

# Look at the value in row 6, column 4
students2014[6, 4]

# Choose the whole column 3 and save it as an object
student_age <- students2014[, 3]

# Choose the third and fourth columns and save them as an object
age_height <- students2014[, 3:4]

# Print out the first five rows of age_height
age_height[1:5, ]



```

*** =sct
```{r}
test_object("age_height")

test_error()
success_msg("Good work!")
```