--- type:NormalExercise lang:r xp:100 skills:1 key:b830d50ea8

## Looping (2)

A loop can be used to fill a vector with values. This is often useful in order to store the results of looping.

One simple way to achieve this is to first create an empty vector using for example `numeric()`. The vector can then be filled inside the loop with new values using indeces.


*** =instructions
- instruction 1
- instruction 2

*** =hint
- hint 1
- hint 2


*** =pre_exercise_code
```{r}
# pre exercise code here
```

*** =sample_code
```{r}
# Size of the experiment
N <- 10

# Create a numeric vector. Print out the values.
vec <- numeric(N)
vec

# fill the vector with new values
for(i in 1:N) {
  vec[i] <- i
}

# print out the new values
vec

```

*** =solution
```{r}
#solution code here
```

*** =sct
```{r}
# submission correctness tests

# example tests:
# test_output_contains("output")
# test_object("object_name")
# test_function("function_name", args=c("arg1"))

# test if the students code produces an error
test_error()

# Final message the student will see upon completing the exercise
success_msg("Good work!")

```


--- type:NormalExercise lang:r xp:100 skills:1
## Brackets have your data [ under, control ]

It is quite often necessary to access only a part of a dataset. This is called *subsetting* and there are many ways of doing it. The best way will depend on the situation and  preference.  

We have already seen that the columns of a data frame (df) can be accessed via `$`. There is another, more general method: `[ ]`. The brackets are a way of accessing a part of any data structure in R. 

The way to use brackets is by giving the locations of the values you wish to access.


*** =instructions
- Study the examples on how to select parts of a data frame.
- Replace `NULL` so that the object `age_height` contains the third and fourth columns of students2014.
- Modify the code to print out the first five rows of `age_height`.


*** =hint
- Selecting multiple columns is very similar to selecting multiple rows.
- See the first example on how to select multiple rows.


*** =pre_exercise_code
```{r}
students2014 <- read.table("http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS-data.txt", sep="\t", header=TRUE)
  
  # keep a couple background variables
  students2014 <- students2014[,c("sukup","toita","ika","pituus","kenka","kone")]
  
  # recode kone -variable NA values as factor levels
  students2014$kone <- addNA(students2014$kone)
  
  # choose rows without missing values
  students2014 <- students2014[complete.cases(students2014),]

```

*** =sample_code
```{r}
# students2014 is available

# Choose the rows from 3 to 6 in the data
students2014[3:6, ]

# Look at the value in row 6, column 4
students2014[6, 4]

# Choose the whole column 3 and save it as an object
student_age <- students2014[, 3]

# Choose the third and fourth columns and save them as an object
age_height <- students2014[, NULL]

# Print out the first five rows of age_height
age_height[NULL, ]


```

*** =solution
```{r}
# students2014 is available

# Choose the rows from 3 to 6 in the data
students2014[3:6, ]

# Look at the value in row 6, column 4
students2014[6, 4]

# Choose the whole column 3 and save it as an object
student_age <- students2014[, 3]

# Choose the third and fourth columns and save them as an object
age_height <- students2014[, 3:4]

# Print out the first five rows of age_height
age_height[1:5, ]



```

*** =sct
```{r}
test_object("age_height")

test_error()
success_msg("Good work!")
```

--- type:NormalExercise lang:r xp:100 skills:1

## Indices and brackets

Vectors store multiple values of the same data type. The values in vectors have indices: the first value in a vector has a index `1`, the second value has `2` and so on. 

You can set or get a single value from a vector by using indices and brackets `[ ]`. Using an index number inside the bracets gives access to a single value from the vector. Using a vector of indices gives access to a vector of values.

It is also possible to rearrange the values in a vector by using indices. Look at the example code to see how indices work. 

*** =instructions
- Create the name vector.
- See how the indices work by executing the example lines.
- Use brackets and indices on `names` to create a new vector `girls` with values "Liisa" and "Anna" (in that order).
- Use brackets and indices on `names` to create a new vector `boys` with values "Pekka" and "Heikki" (in that order)

*** =hint
- Note that space between the vector object and bracets produces an error.
- Remember that vectors are created with the function `c()`.
- Index vectors `c(1,2)` and `c(2,1)` do not produce the same outcome. The order of the values is different.

*** =pre_exercise_code
```{r}
```


*** =sample_code
```{r}
# Let's create a vector
names <- c("Matti", "Pekka", "Liisa", "Anna")

# Acess the first value of the vector
names[1]

# Change the first value of the vector
names[1] <- "Heikki"

# Acess the 1. and 3. value of the vector
names[c(1, 3)]

# Use indices and brackets to separate the names vector into two vectors of the specified ordering
girls <-
boys <-

```

*** =solution
```{r}
# Let's create a vector
names <- c("Matti", "Pekka", "Liisa", "Anna")

# Acess the first value of the vector
names[1]

# Change the first value of the vector
names[1] <- "Heikki"

# Acess the 1. and 3. value of the vector
names[c(1, 3)]

# Use indices and brackets to separate the names vector into two vectors of the specified ordering
girls <- names[c(3, 4)]
boys <- names[c(2, 1)]

```

*** =sct
```{r}

test_object("girls", incorrect_msg = "Did you create `girls`?")
test_object("boys", incorrect_msg = "Did you create `boys`? Are the values in the correct order?")

test_error()

# Final message the student will see upon completing the exercise
success_msg("Such indices. Wow. Much intelligent.")

```


--- type:NormalExercise lang:r xp:100 skills:1
## Data types and functions

Not all data types do well with certain functions. 

One cannot for example use character vectors when the function needs a numeric one. It is also impossiple to create a vector with multiple data types. 

When using the wrong data type, R produces an error stating what is wrong. 

*** =instructions
- Create `my_vector`.
- Try to calculate the mean of `my_vector` and see what happens. 
- Correct the vector and calculate the mean so that you do not get any errors.

*** =hint
- You can use `str()` to check the data types.
- `chr` is short for character.
- When creating a character vector, one needs to use quotation marks. 

*** =pre_exercise_code
```{r}
```

*** =sample_code
```{r}
# Let's create a vector
my_vector <- c(2, 3, "1", 5)

# Calculate the mean of vector
mean(my_vector)
  
```

*** =solution
```{r}
# Let's create a vector
my_vector <- c(2, 3, 1, 5)

# Calculate the mean of vector
mean(my_vector)

```

*** =sct
```{r}
test_object("my_vector", incorrect_msg = "Did you correct the `my_vector`?")

test_error()

# Final message the student will see upon completing the exercise
success_msg("Creating vectors like a pro. Great work!")

```
