Suunnitelma DataCampin osioille 6, 7, 8, 9  

## 6. Subsets and conditioning

### 6.1 Selecting a subset
Introduction to using subset()

- newdata <- subset(data, conditions)

### 6.2 Logical conditions
Introduce the R logical conditions

operator  | description
----------|------------
`==`      | exactly equal to
`!=`      | not equal to
`<`       | less than
`>=`      | greater or equal to

### 6.3 Logical operators

operator  | description
----------|------------
! x       | NOT x
`x & y`   | x AND y
`x | y`   | x OR y

-  more subsetting

### 6.2 Long term frequency
Use sample() to demonstrate the long term frequency of some condition as a function of the sample size. 

- Example: the frequency of succesfully choosing a male student from the data
- Compare to prop.table() ?


### 6.4 Proportions and probability
Use the logical operators to answer probability questions

- Example: what is the probability the a randomly selected students height is less than 160
- nrow() 

### 6.5 Conditioning
Use the logical operators to answer more difficult probability questions and demonstrate conditioning

- What is the probability that a randomly selected female student has a windows computer?

## 7. Probability distributions


### 7.1 Distributions: continuous or discrete?

- multiplechoice?
- precode?/console?: plots of continuous/discrete plots

- perhaps move this to the end? (tuomo)

### 7.2 Uniform distribution

- Could use sample() to demonstrate discrete uniform distribution and/or create own probability distribution.
- Example: 

```{r}
tomorrow <- c("cat attacks baby","world ends","I eat pasta")
probabilities <- c(0.2, 0.3, 0.5)
# draw a sample
sample(tomorrow, size = 1, prob = probabilities)
```

```{r}
# draw a sample 100 times and visualize the distribution
x <- sample(tomorrow, size = 100, replace = T, prob = probabilities)
barplot(table(x))
```


### 7.3 Binomial distribution

### 7.4 Normal distribution

### 7.5 Create your own functions

### 7.6 Standardization

- create function to standardize
- normitus <- function(x) x - mean(x) / sd(x)


## 8. Simulations and confidence intervals


### 8.1 Advanced vectors
Introduce the bracket notation and indexing 

- 1:10 notation
- Brackets and indexing

### 8.2 Looping
Simple demonstration of the for loop

- Simple exercise utilizing a loop and filling a vector

### 8.3 Bootstrapping
Leikitään, että students2014 edustaa perusjoukkoa

- Samplataan datasta 30 oppilasta lasketaan (esim) pituuden keskiarvo. toistetaan useita kertoja ja tallennetaan tulokset
- Tuotetaan odotusarvon estimaattorin empiirinen jakauma ja tarkastellaan sitä. histogram(), summary()

### 8.4 Central limit theorem
Yleistetään edellisen tehtävän huomio tarkastelemalla jonkun toisen muuttujan keskiarvon empiiristä jakaumaa

- Huomioidaan millä välillä suurin osa arvoista (95%) on

### 8.5 A Confidence interval
Otetaan yksi otos ja lasketaan analyyttisesti 95% luottamusväli

- sama parametri kuin edellisessä tehtävässä. 
- Vertaillaan empiirisiin tuloksiin.

### 8.6 Confidence interval demo
Otetaan nyt monta otosta (for loop) ja lasketaan luottamusvälit jokaisen kohdalla. Tallennetaan ne.

- Piirretään kuva josta nähdään kuinka suuri osuus luottamusväleistä sisältää todellisen parametrin arvon (samaan tyyliin kuin https://tuomonieminen.shinyapps.io/CIdemo/)
- Tämä on varmaankin enemmänkin demonstraatio kuin tehtävä

## 9. Hypothesis testing

### 9.1 Statistical hypothesis testing
- basic consepts

### 9.2 Meet the tests!
- maybe more than one exercise...
- t/z/p-tests, t-distribution

### 9.3 Peculiar p-values

### 9.4 (Null hypothesis:) which way to go?!
